### 2022-05-03 gkw/doc/etc/devel.tex

%% GKW developer notes: presently only rules and guidelines for developers. 
%
\documentclass[12pt,twoside,notitlepage,a4paper]{article}

%% caption modifications; default to footnotesize - tweak as appropriate
\usepackage[font={footnotesize,it}]{caption}
%\usepackage[small,labelfont=bf,textfont=it,width=0.9\textwidth]{caption}

%% for modifying description lists 
%\usepackage{expdlist}

%% the listings package; useful for blocks of code
\usepackage{listings}

%% fancy verbatim {Verbatim}
\usepackage{fancyvrb}

%% for using some regular formatting in verbatim {alltt}
%\usepackage{alltt}

%% underscore
\usepackage{underscore}

%% graphics
%\usepackage{graphicx}

%% look for figure input in ./figs
%%\graphicspath{{../figs/}}

%% fullpage
\usepackage{fullpage}

%% color
\usepackage{color}

%% makeindex ?
%\usepackage{makeidx}

\usepackage{bm}

%% hyperref
\usepackage[hyperindex]{hyperref}

% ---------------------------------------------------------------------------- %
%% Macros
\newcommand{\file}[1]{\texttt{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\variable}[1]{\texttt{#1}}

%%\newenvironment{mylisting}
%%{\begin{list}{}{\setlength{\leftmargin}{1em}}\item\scriptsize\bfseries}
%%{\end{list}}
%%
%%\newenvironment{mytinylisting}
%%{\begin{list}{}{\setlength{\leftmargin}{1em}}\item\tiny\bfseries}
%%{\end{list}}

% ---------------------------------------------------------------------------- %

\begin{document}
\pagestyle{empty}
\pagestyle{plain}
\pagenumbering{roman}

% ============================================================================ %

\begin{titlepage}
  \begin{center}

  \large

  \vspace*{2cm}

  {\Large\sffamily\bfseries GKW:\\[2\parskip]
    Notes}

  \vspace{3ex}

  {\sffamily DeveloperGuidelines}

  \vspace{\stretch{2}}

  \vspace{\stretch{2}}

  \vspace{\stretch{1}}

  \emph{$ $Date$ $, $ $Revision$ $}\\
  \url{http://gkw.googlecode.com}

  \vspace{\stretch{3}}

\end{center}

\end{titlepage}

\clearpage
\tableofcontents
\cleardoublepage
%\pagestyle{fancy}
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\cleardoublepage
\appendix
\part{Appendix}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% ============================================================================ %

\section{Coding GKW: rules, style and recommendations}

It is intended that this appendix becomes a part of the main code documentation.
Its purpose should be to make clear what standards are expected of code or files
committed to the repository.
For now, we should use it as a discussion about what exactly these things should be.
The motivations for adopting such conventions are fairly clear (e.g. keeping the code
readable, portable, maintainable, flexible etc.). We should aim to
find a {\it reasonable} set of conventions to agree on, rather than iterate to perfection.

I suggest that we initially focus only on the Fortran source parts of the code, then
develop conventions for other files later as necessary. It may also be worth adding
a section of notes about ways of avoiding certain undesirable features (e.g.
non-standard/extended Fortran -- just a list of things we already know about that you should not
do).

So far we have adopted some conventions and style fairly consistently throughout the code
and these can be documented here already. There are some conventions which should be
clarified and perhaps more that should be adopted. If the code is tidied a bit, some
conventions will become self-evident and we may not need to document them here.

Please discuss (in this document file even). The following could be restructured a bit...

FC, YC:  I agree with all, except where I have commented otherwise.\\
{\bf YC: it looks like GKW is not in a state consistent with the guidelines. How do we implement the changes ?
Each of us takes a couple of files and does it ? APS: it is not difficult for one of use to do it. Some checks
are difficult so I will likely write a short script to do them. It might be
worth 2 people first independently look look at 1 module, then review what the other did to see
if it meets their understanding of the conventions...}

\subsection{General}

The code is written in standard Fortran 95. Extensions to the standard are not
allowed and redundant features should be avoided. The code should always
be able to compile and run on any system which has a working Fortran compiler.

External Fortran libraries may be added to the repository to provide extra     
functionality to the code. These should be no less portable than the rest
of the code and their license must be compatible with the one with which the
code is distributed. Anything copied from some other source should have its
origin made very clear, and again, do make sure that it is allowed to
be distributed together with the rest of the code (If in doubt, please
discuss with the code administrators.). 
%APS:
{\bf APS: Is this OK? If so, must the added code meet any conventions? For example, 
our FFT interface follows different style and conventions to the rest of the code.
Also, can any added code have unintended consequences? What should the policy be?}

{\bf AGP: Everything in /trunk/src should follow the style rules, even if we 
stole it from somewhere else. Is there a good suggestion on where to store library 
source code. APS: separate subdirectory at the same level as trunk, branches; 
README file making it clear it is not part
of the code; named `external', `lib', ? (I don't like any of these names very  much). }

The code may make use of calls to other 3rd-party libraries, provided they have
some sort of Fortran interface. Calls to such libraries should be hidden away
from the rest of the code as much as possible. The use of non-portable and
non-free software libraries is strongly discouraged; they should never be
enabled as default and must never constitute an essential code component;
they should only be used to enhance performance of the code on specific systems
rather than functionality. (FC: The OpenMP in exp integration currently makes use of the Intel MKL, 
but as discussed in issue 5 this is only a temporary arrangment during testing and development.)

We aim to make the code easy to read, understand, maintain and modify. 
To this end, we make many functions and subroutines as general as possible to encourage code re-use.
The following sections highlight particular aspects of this style and adopted conventions.

Suggestions regarding adding new features to the code, restructuring the existing code etc. can be discussed in the discussion list.

\subsection{Files}

Each source code file should contain either a Fortran program or a module.
The name of the file should correspond to the program or module contained within it,
followed by either a \file{.f90} or \file{.F90} suffix (the \file{.F90} suffix
denoting a file which must be preprocessed -- see {\it preprocessing} later).

Including files via the obsolete \code{include} statement should be avoided.
An exception is made for the interfaces to Fourier transforms and MPI libraries etc. -- if a file
must be included it should be included only once in a single file.

\subsection{Names}

Modules, subroutines, functions, variables etc. should be given meaningful or self-explanatory
names in general (be careful with spellings!). The names should be fairly concise, but not so short
that they are not understood and absolutely not more than 32 characters. For indices that are used in loops, the variable names
need only a single letter if desired (and should probably be kept short?). {\bf FC: I think two letters max is plenty.
APS: so we should explicitly discourage/disallow long indices names?}
{\bf AGP: to letters max is not a bad idea, but the code does violate this restriction rather strongly 
at present (imod). I think furthermore that we should encourage the use of standard indices for certain 
loops (for instance imod (or better im) for the toroidal mode number}
\textbf{YC: I do not really care about the length of the indice names as long as it is clear from their name that they are indices (which is I think is the case for imod) and are only used to that purpose. }
All names should be written in lower case in the code source files, apart from a few exceptions

\begin{itemize}
\item Optional function and subroutine Keywords.
\item External variable and subroutine names (e.g. the integer \code{MPI_OFFSET_KIND} is a
MPI variable.)
\item
%APS: 
{\bf APS: Anything else? We could add macro definitions, directives etc? (I vote for these)}
\end{itemize}

\subsection{Modules}

Each module should have a clearly defined purpose or set of tasks associated
with it as described in its doxygen compatible leading comment. Usually, all functions, subroutines and variables should be
private to that module my default and only those used elsewhere should be
exposed via the \code{public} statement.

\subsection{The use of \code{use} statements}

Every \code{use} statement should always be followed by the \code{only} option to explicitly
state which variables and functions are in scope. There is one exception for the MPI
interfaces, i.e. every module which has a name beginning `mpi' may drop this convention.
%APS:
{\bf APS: this need not be the case, only at the moment you would need to put all the MPI
things in too, e.g.
\begin{verbatim}
  use mpiinterface, only : MPI_ALLREDUCE, MPI_TYPE_CREATE_SUBARRAY
\end{verbatim}
etc.}

%Since the routines in a module typically work on the same arrays and structures etc.,
%a commonly used convention is to declare all the variables and routines used in
%other modules once globally at the module level via USE statements.
Rather than putting all the required \code{use} statements for a module at the beginning,
we have adopted a convention that \code{use} statements should only appear when required -- i.e.
they should pertain only to the subroutine in which they exist. 
%APS:
{\bf APS: this does not preclude the possibility of them appearing at the beginning of a module,
since they well may be necessary to define variables that are global to the module.}
{\bf AGP: if really necessary for the definition of global variables they may appear at the 
top, but only then. (This is not done consistently in the code at present)}
This enables one to see exactly what is within scope at the subroutine level.
To maximise the usefulness of this convention, only those variables
actually used should be imported to each routine. When starting a new routine, it
is very tempting to copy the \code{use} statements from a neighbouring code block without
checking if all variables are actually needed. When doing so, you should check that
any variable not used is removed.

\subsection{Interfaces}

%APS:
{\bf APS: There is a suggestion that we should restrict function overloading
to functions which have the same number of arguments. For example, the
\code{indx} function gets called with either 4 or 6 arguments, so perhaps
there should be two distinct functions that are called. Other overloads, such
as \code{mpibcast}, are fine as they always take the same number of arguments,
independent of datatype.}
{\bf FC:I see no benefit in overloading functions through interfaces, it only makes 
the actual bit of code take longer to find and makes the code harder to read. APS:
The purpose of interfaces is to make things easier all round. If something is not clear
we should change it. I don't believe the interfaces themselves are the problem.
Say I want to do a MPI_BCAST. Then doing it explicitly I have to (i) look up the interface
\url{http://www.mcs.anl.gov/research/projects/mpi/www/www3/MPI_Bcast.html}; (ii) include
preprocessor directives. (iii) check all the datatypes are correct. Using the 
overloaded function I get to know at compile time if I've made a mistake (rather than when the code
runs with some obscure MPI error message). (and the
wrapper can be used to handle all the mpi things in a consistent way rather than having to
repeat the same code over and over elsewhere).
FC: I also suggest generally avoiding optional arguments, they also make the code harder to read. APS:
sometimes we would be left with passing many lines of unnecessary variables to
a subroutine -- it would be nice to see which ones are actually being used. I agree with the examples
we discussed earlier, but these should be indentified in the
code and made clear.}

\subsection{Some things not allowed}

\begin{itemize}
\item No variable may have the pointer attribute. We do not use pointers in the code.
\item Derived data types should be avoided for the most numeric parts of the code (and in general?)
where potential compiler optimisation may be lost.
\end{itemize}

\subsection{Variable declaration}

\begin{itemize}
%APS:
\item {APS: \bf should we use \code{dimension} or not? I think putting the dimension after the
variable instead is fine, but we should stick to one or the other.} FC: Agree, no strong preference.
\item We do not use non-standard \code{real*8} etc. Instead we use \code{character (len=32)}, 
\code{real (kind=somekind)}, etc.
\item We use \code{::} for all declarations, not only for consistency but to avoid problems with some
compilers and also clarity.
\item Every dummy variable in a subroutine should have an \code{intent} (the default is
{\it undefined!}). This includes intent \code{inout}.
\end{itemize}

\subsection{Subroutines}

Individual subroutines should be kept short, not more than say $\sim200$ lines. A subroutine likely
tries to do too much if it is long and could be split into more subroutines.
%APS:
{\bf APS: perhaps part of the reason we are not encouraged to break subroutines is that we have
to set up all the use statements again? Subroutines can themselves contain subroutines, which
gets around this.}

All optional arguments should be capitalised.
e.g. the subroutine
\begin{verbatim}
subroutine mpibcast_real_array1(rscalar,nscalar,IPROC)

  real, intent(in) :: rscalar
  integer, intent(in) :: nscalar
  integer, optional, intent(in) :: IPROC
\end{verbatim}
with optional argument \code{IPROC} and its call
\begin{verbatim}
  call mpibcast(some_real,some_integer,IPROC=some_proc)
\end{verbatim}

\subsection{Code comments}

Comments are important! We can certainly have too many comments. We should aim
to explain what each routine does in the description. No need to go overboard;
comment should say what task is being performed; "This is a loop over x..."
is useless?

This is bad? (and hard to read?)
\begin{verbatim}
  ! something
  call do_something()
  ! something else
  call do_something_else()
  ! more still
  call do_more()
\end{verbatim}
is this better?
\begin{verbatim}
  ! something
  call do_something()

  ! something else
  call do_something_else()

  ! more still
  call do_more()
\end{verbatim}

Comments in GKW should be written in such a way as to be parsable by \href{http://www.doxygen.org}{Doxygen}.
The \href{http://www.gkw.org.uk/doxygen}{reference dictionary} for the code can then be automatically generated.

\begin{itemize}
\item All modules, functions and (module level) variable declarations have a useful Doxygen comment.
{\bf APS: should it be public variables, rather than module level?}
{\bf AGP: I would say module level as well, since a module can 5000 lines, and rather confusig to look through}
\item Make informative comments thinking of others {\bf APS: Sentences in English are very much appreciated!}
\item Notes to self should be marked with your initials e.g. \texttt{!APS: check this} {\bf APS: These should be in col 1 and considered as bugs! we should aim
to remove them. Note should be meaningful to all, not just creator; should it be dated?}
\textbf{YC: yes, I like the idea of dating it. It makes it easier to guess what is right when two comments are inconsistent.}
\item Questions to others should be marked e.g \texttt{!APS -> WAH: is this correct?} {\bf APS: I dont think it is so useful. Issue tracker is fine. initials 
considered as bugs; always start in the first col. Also all preprocessor and compiler directives. I think all other comments need a space between \code{!} and the first real
character of the comment.}
\item Input variables are labelled as such with default values listed if applicable at declaration comment
\item Allocatable arrays should have their size noted at declaration comment 
\item Indent comments inline with code
\item How do we do block comments?
\item Commented out code starts \code{!NAME: }? or what? Double \code{!}, with a
tag to explain who commented it out and why?
\textbf{YC: yes, I think it is good practice to explain why part of the code is commented out. Otherwise, why not remove it?}
%APS: \item Unused code should be commented out? Then comment rules apply...
%APS: comments are bugs?
\end{itemize}

\subsubsection{How to write comments compatible with the doxygen parser}

\textbf{Variables} can have either a leading comment, or trailing comments on same line.

\begin{itemize}
\item 
\textbf{Example 1:}\end{itemize}

\begin{verbatim}
!> the total energy flux per species : eflux_tot(number_of_species) 
!> electrostatic contribution 
real, allocatable, dimension(:) :: eflux_tot_es 
\end{verbatim}

\begin{itemize}
\item 
\textbf{Example 2:}\end{itemize}

\begin{verbatim}
logical :: coriolis             !< true if coriolis drift term is kept 
\end{verbatim}

\textbf{Functions}, \textbf{subroutines} and \textbf{modules} must have their description \textbf{preceding} the declaration. To keep it simple, use the \texttt{!>} comment format for all comment lines above a function declaration.  Lines will run together as a single sentence.

\begin{itemize}
\item 
\textbf{Example 3:}\end{itemize}

\begin{verbatim}
!-----------------------------------------------------------------------------
!> This routine calculates the correction necessary to go from the
!> distribution g (which includes the correction of the parallel
!> vector potential) to the distribution f
!-----------------------------------------------------------------------------
subroutine g2f_correct
..
..
\end{verbatim}

Alternatively, comments preceding a function can be written as a formatted block which will retain its layout through the parser (in next example).

Latex math mode can be used if enclosed within two \texttt{\\f\$} groups.  If used, it must
be valid latex without macros or packages - or the doxygen parser will fail.  See
\texttt{linear\_terms} for examples.

\begin{itemize}
\item 
\textbf{Example 4:}\end{itemize}

\begin{verbatim}
!++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
!> Add terms I and IV as one term : \f$ v_R {\cal F} \{H,g_N\} \f$
!! where
!!         \f$ H  = (1/2) v_{\parallel N}^2 + \mu_N B_N + (1/2) {\cal E}_R = H(s,\mu,v_{\parallel}) \f$,
!! and
!!         \f$ \{H,g_N\} = (d H/d s)(d g_N/d v) - (d H/d v)(d g_N/d s). \f$
!!
!! At the boundaries in the s-direction, we apply something similar to the
!! wills=1 scheme to do upwinding. Note that the routines that this calls to
!! deal with points near the boundaries do not presently work in second order.
!!
!! See jhg_interior for the main bit of these terms.
!!
!! ( Note that above we dropped the second index with respect to `j'(for \mu)
!!  in H; the second index is always j. Below we use `HH' for `H')
!<----------------------------------------------------------------------------
subroutine igh(disp_par,disp_vp)
...
... 
\end{verbatim}
{\bf APS: in the last two examples above, is the \code{-} character the one we
have decided to use to break up the code blocks in all cases? Only in the code we use at least \code{+,*}
as well...}
\subsection{Code blocks, line indentation, whitespace and continuation}

We do not allow TAB characters in the code. They may mean something in an extension of Fortran, so
could have unintended consequences. Also, they can make reading the source between different
editors difficult.

Blank lines may appear in the code (see comments section); they should contain no characters (including
whitespace characters). A line may be a blank comment line (e.g. in a comment block), but should
start with \code{!}. FC: Comments should indented inline with code.

\begin{itemize}
\item \code{module, subroutine, function, contains statements etc. begin in col 1} 
(unless they are functions inside subroutines?)
\item code is indented 2 columns from the program unit.  FC: I prefer 3. AGP: I prefer 2. YC: 2 for me 
\item \code{do,if,select case} require indent of 2 cols.  FC: I prefer 3. AGP: I prefer 2. YC: 2 for me. APS: so it is 0 or 2 and never 3?
\item Maximum line length. There is no need to use the old Fortran 77 limit. However, anyone
editing with emacs or vi would greatly appreciate it if most lines do not go beyond 78 cols
(I suspect most potential coders would use one editor or the other). What lines should go beyond the
limit? Lines that do not need to be read to understand what the code does? FC: I vote for more than 78.
APS: The original code used around the Fortran 77 limit. Since the style should be similar, we need not
usually use more than that. If we set a _hard_ limit around 78, and a suggest using less in general, that
already allows one to use several more cols. If we agree that this is too restrictive, why not set the hard
limit to 132 (as allowed by the standard). To make it clear - the main reason for 78 is that when remote
editing a file (such as when you do some tweaking to the source you are compiling on a cluster), it is usual
and convenient to run your editor in non-graphical mode. When doing so, lines longer than 78 are wrapped in
emacs and it can be awkward to work with. \textbf{YC: I would prefer the limit to be 132, but it is because I never run an editor in non-graphical mode...}
\item Long lines should be broken by placing a continuation character (\code{\&}) in the last col?
A continuation character should also appear in the next line, indented 4(?) cols in from the start
of the previous line. Lines which end part way through a character string should be terminated before
the continuation character, then concatenated with another string on the next line.
e.g. {\bf AGP: continuation characters should be placed in a readable place, for instance directly 
under the = in an assignment statement.APS: In the line below it could instead go under the opening brace?
Setting to a fixed number is easy for editors though - I don't have a macro to
automatically put it in the right place}
\begin{Verbatim}
call gkw_init('gkw_init: I do something related to reading files and'//      &
    &         ' allocating memory')
\end{Verbatim}
\end{itemize}

{\bf Need we discuss where whitespace is used in e.g. \code{if, do}?}

\subsection{Misc - unsorted ideas}

\begin{itemize}
\item \code{endif,enddo,endsubroutine} vs \code{end if, end do, end subroutine}; should we be consistent? 
AGP: Not a major point for me. APS: looking in the code, mostly the words are
split, apart from "endif" (roughly 4 times more common than "end if").
\item Should we use the operators $<, >, <=, >=, ==, /=$? They are easy to read and the
old ones are considered obsolete. AGP: Yes would be nice 
\item Should we avoid using multiple \code{return} statements? They are redundant at the end of subroutines
anyhow. Sometimes it is difficult to see which code gets reached if there are many. Any thoughts?
The only advantage is in an optimised routine when many more unnecessary tasks would otherwise be
performed (but these should be in sub, or subsub routines, right?) AGP: I have always considered multiple 
returns useful. Often there are special cases in which the job gets done with just a few lines. APS: certainly in
some optimal cases. Not needed at the end of subroutines or functions though. I'll refer other specific cases to you.
Some are really difficult to debug.
\item Should we use \code{()} for function and subroutine calls which have no variables to pass? FC: yes.
\item Numbered lines - are they needed? Formats in write statements are the only thing that use them, but
even that is not necessary. FC: neater and easier for (repeated) write statments. APS: Rather than have
numbered lines all over the place, I suggest using character strings instead. Even better, putting
all the formats in a module, so that they are done consistently throughout (and could potentially be
modified via a namelist). AGP: indeed numbered lines only for the write statements. Putting them in 
a module is not a bad idea. 
\item Do we allow \code{;} to e.g. put multiple assignments on one line or for multiple do loops etc.?
When and why?  FC: yes as it keeps routines shorter and more readable. AGP: After being initially against
it, I now have changed my mind. Indeed it does make the code a lot more readable. So yes. 
\item No hard coded numbers? - should use \code{parameter} instead?  FC: Clarify $10^-10$ not ``two''.  Does this include: \texttt{max_iterations}, \texttt{nmmx}? AGP : I prefer no hard coded numbers (although I am guilty) YC: what do we do for things like $2\pi$, etc.? We want to have it with the same precision everywhere in the code, no ?
\item Always use \code{implicit none} exactly once in each module. 
\item Array syntax. It should be kept simple in the optimal parts of the code, right? FC: Don't understand. APS: OK, when should we use (if ever) array syntax?
It is extremely concise and usually at least as efficient as doing it explicitly, but mistakes can be made if you try to be a bit tricky.
(when assigning with array syntax, use \code{A(:) = 0.}, rather than \code{A = 0.}? etc.) AGP : I agree use A(:) and preferably no array syntax a = b + c. It can be confusing
\end{itemize}

\subsection{Input files}

Input files are also Fortran files! (Only applies to files checked into the repository, such
as the test files and samples.)
(and we have had some problems with them! We use some non-standard features which could be avoided
by introducing simple guidelines for writing them.)
How about:

\begin{Verbatim}[commandchars=\\\{\}]
! This is a sample input file.
! Comments are allowed as in normal fortran while outside a namelist block.
! No non-Fortran chars allowed! (Including TABS)
       \textit{! Blank lines are allowed}

&CONTROL  \textit{! no need to be capitals; comments allowed this line; \code{\&} in col 1 only}
 some_variable = 1. \textit{! variables indented 1 space, comments only on same line}
 another_variable = 1, ! comment allowed here
       ! comment allowed here (unless the previous variable is a character, if that
	   ! is the correct interpretation of the standard) NO TABS!
 \textit{(some compilers apparently do not allow comments in the block at all;}
 \textit{ not sure the standard allows them on next line unless they start}
 \textit{col 1 and the previous line has no comma at the end..)}

/ \textit{! end of block starts in col 1}
\end{Verbatim}

in summary:
\begin{itemize}
\item Blocks start and end in col. 1
\item \code{var = val} starts in col. 2
\item No TABS
\item No comments after \code{char=val}. The comment can go on the next line or previous line.
\item If the line ends with \code{","} with no comment afterwards, the next line can not be a comment?
\item Stick to the same line length as for the fortran files (when we decide)
\end{itemize}

\subsection{Preprocessing}

We use C preprocessor directives in the code for the purpose of conditional compilation only.
They are used to enable/disable calls to optional libraries, such as MPI, FFTW, UMFPACK etc.
and also to get the right datatypes in those calls. Ideally, the actual calls to
these external libraries should limited to 1 module per library (this is already the
case for FFTW, and almost true for the MPI).

The compiler should see as much of the code as possible when testing, so we avoid
any other use. (One can use the timings module \file{PERF}, together with directives
for testing parts of the code, but putting these all over the checked in code make it messy?
Perhaps we could always call perf at the desired places, but either put a 
logical "perf" (which decides if the call actually does anything), or preprocess perf itself
with the list of perf names)

Allowed list:
\begin{verbatim}
#ifdef ONE || TWO
#elif defined(THREE)
#else
#endif
\end{verbatim}
(Do we need more?)
\subsection{Fortran best practices}

Do we need such a section? Most people know it is bad to put conditionals in the middle of 
optimised loops, that all cases should be covered by conditionals etc. {\bf APS: I have a list.
}. \textbf{YC: This would be useful to me. The little I know in Fortran is from copy-pasting bits of code. So I might very well have taken bad practices with it.
APS: Right, the list can begin with things we removed already.
Then whenever cleaning up, add more things to the list.
The old features can be summarised in another section again, although they can be found in most fortran books.}. We should always try to generalise bits of code though to encourage re-use, so
examples of generalisation could go here. Obsolete and old features that are not used in the code
could also be listed.
{\bf AGP : Optimized routines should be flagged as such. It is often very useful to program 
with readability in mind rather than performance. This should be encouraged in all the 
non-optimized routines. Only a few routines of the code need really be optimized for speed}

\subsection{Summary of main unresolved points}

\begin{description}
 \item Most things that were here are resolved or unimportant.
%  \item If \code{dimension} should be used or not.
%  \item What character(s) should be used to break up code blocks in comment lines.
%Presently, at least \code{-,*} and \code{+} are used.
%  \item The number of indentation whitespace characters used. We agreed on 2 in all cases, apart from the
%  start of a subroutine (no indent, or 2)? In the special case of continuation, is it always clear where
%  to put it on the next line? it could be omitted completely. Do we indent the code under each \code{case} in a \code{select case}?
%  \item The first continuation character should go at the end of the line, or not? (i.e. if we allow 78 cols it goes in col 78)
%  \item The number of allowed cols. 78 vs more?
  \item \code{endif} vs \code{end if}: see A.12
\end{description}

%\section{The previous section again}

%%%% ============================================================================ %
%%%% Index 
%%%% Put the indexes here for files, variables, then another for names,
%%%% defs, commands etc.
%%%%
%%%% ============================================================================ %

\end{document}