### 2022-05-03 gkw/doc/manual/buildandrun.tex

\chapter{Obtaining, building and running the code} \label{usage}

\section{Getting the code}

To obtain the latest version of the code (which should approximately
correspond to the latest version of this documentation), you should use the
git revision control system (\href{http://git-scm.com/.}
{http://git-scm.com/.}). The git program is installed on many
systems as the \name{git} command. Assuming you are using \name{git}, you can
check out the code with the command
%git clone http://gkw@bitbucket.org/gkw/gkw.git gkw-read-only
\begin{verbatim}
  git clone git@bitbucket.org:gkw/gkw.git gkw
\end{verbatim}
To push changes, you will need a bitbucket account to be given write permissions.

% Not yet sure how this will be relvant to bitbucket:
%[Some advice on which tagged versions to use, or how to download without git.]%

% This clone does not allow changes to be written back to the repository.
% Instructions for developers with commit permissions 
% can be found at \href{http://code.google.com/p/gkw/source/checkout}{
% http://code.google.com/p/gkw/source/checkout}.

% Alternatively, a reasonably stable version of the code can be obtained as a
% tarball from the downloads section
% (\href{http://code.google.com/p/gkw/downloads/list}
% {http://code.google.com/p/gkw/downloads/list}).
% At present, the version of the code found there is based on the CPC release
% (associated with the CPC paper \cite{CPC-paper}). The tarball is not often
% updated and will not have the latest code features.

\section{Building the source code}

% If using a version obtained in a tarball, it is suggested to look at the
% \File{README} file found within for specific compiling instructions. The
% guidance of the following sections pertains primarily to the latest version of
% the code obtained via Subversion.

There are two methods provided to build the code, involving various
makefiles. The use of the `simple' makefile, \File{src/makefile}, is intended
only as a fallback if the standard makefiles, requiring \name{GNU make},
cannot be used.  The standard GNU makefiles
allow one to create build settings on a per user/machine/compiler basis,
without having to make any modifications to the original makefiles\footnote{
The standard makefiles also result in executables which can
provide information about themselves and how they were produced. For example,
the executable will be able to report its revision number in various ways,
which can be quite useful to know.}.
Both methods are outlined below.

With the latest version of the source code, it will often be possible to
build the code by doing very little other than running \name{make} in
the top-level directory containing the \File{GNUmakefile} file.
Although this may appear successful, one should be aware that the resulting
executable may not contain all the desired functionality.

\subsection{Prerequisites and suggestions}

A Fortran 95 compiler is essential for building the code. Most Fortran
compilers are able to pre-process the source files containing C
pre-processor directives\footnote{In the unlikely event that this is not the case,
and an appropriate pre-processor is not available, some editing of the
source files may be necessary before compilation.}.

A minimum of a Unix-like system, together with the POSIX shell and standard
\name{make} utility is suggested.\footnote{Without these basics, the various source files will have to be
compiled in the correct order (or concatenated into a single file in the the
correct compilation order beforehand).}
%We are typically developing the code under
%GNU/Linux, which meets this minimum requirement -- as do many other operating
%systems.

For building the latest version of the code from the git repository,
\href{http://www.gnu.org./software/make/}{\name{GNU make}} is recommended.
This should be the default
\name{make} program installed on GNU operating systems, such as GNU/Linux.
On some systems, the command may be \name{gmake}, in order to distinguish it
from the system default \name{make}.

To enable parallelisation, you will require an MPI implementation and/or a
Fortran compiler which supports OpenMP directives.  To allow nonlinear runs, 
the FFTW3 library is required (more details in \Sec{thecode}).

A C compiler is {\it not} required to compile the code, but will be required
to compile the optional libraries includes with GKW.

More information on linking optional libraries is given in 
section \ref{sec:additional-libraries}.

\subsection{Using the GNU makefile}

Instructions within the top-level makefile, \File{GNUmakefile}, should be the
first point of reference for this method. Ultimately, you should just need to
run \name{make} (GNU make) in this directory and you will end up with an
executable in the \File{run/} directory. However, when building for the first
time on a given machine, it is suggested to create at least one new makefile
which contains settings specific to the local machine and user (and perhaps
compiler if there are multiple possibilities). Whenever the code is then built
again on this machine, those specific makefiles will be used
automatically when running make.

All host-specific makefiles are located in the \File{config/} directory and
subdirectories. The variables contained in \File{global_defaults.mk} will
always be used, unless host-specific variables override them. You can use
\File{template.mk}, which documents almost all the variables one may wish
to change. If you are creating a new file for machine \name{hostname}, it
should usually be saved to \File{defaults.mk} within the \File{hostname/}
subdirectory of \File{config/}. You will likely want to set at least the
Fortran compiler, external libraries, and the required precision. One may
wish to look at existing files for examples; template files for Intel, GNU
and IBM compilers are also provided in \File{config/templates}.

The code object files are built in different subdirectories of \File{/obj}, 
depending on hostname, compiler and precision, among other things. The
resulting executable name will represent which git
version of the code was used, which compiler was used, and if the
executable is single (SP) or double precision (DP)\footnote{Beyond these 
distinctions, it is possible to clobber an
existing executable with one of the same name that does something different.
%(this shortcoming may change in the future).
}. 
If you run \texttt{make clean}, ALL object files will removed.
\texttt{make cleanloc} will remove only object files for the compiler and
 precision subfolder\footnote{Note that if you pass variables to
the make program by specifying them after the command (rather than
changing them in the appropriate file), you should be aware that existing
objects might be re-used to complete the build, such that the desired
properties of the program are only partly present. For example, say you
compile the code without OpenMP, then edit some files, then type
\texttt{make SMP=OPENMP}, then it might be the case that only some of the
code has OpenMP enabled in it. To avoid such problems, use the configuration
files, or type \texttt{make clean}, before
 (re)building to make sure you get what you expect.}. 
If your make program supports it, there should be no problem in
building the code in parallel (GNU make does this via the `-j' flag).

If everything completes correctly, the compiled executable will appear in the 
\File{/run} directory.

%\pagebreak

\subsection{Using the `simple' makefile}

If using the GNU makefiles is problematic, one may wish to build the code
using the simple makefile, \File{src/makefile}. You will need to run
\name{make} from within the \File{src/} directory, but there is a good chance
you will have to edit the makefile beforehand -- instructions for doing this
can be found in the file itself. Any files created during compilation will be
in this directory, together with the source files, apart from the resulting
executable, which by default, will be found in the top-level
directory.

\subsection{Some important makefiles}

There are two files common to both methods which are found in the \File{src/}
directory. The \File{objfiles.mk} file contains the list of object files to be
created before linking to produce the executable. It should only be necessary
to modify this file when adding new source files to the code.
The file \File{deps.mk}
lists the dependencies for the objects found in \File{objfiles.mk}, so that
the objects/modules may be built in the correct order. This should be updated
automatically with the standard makefiles, but may need to be modified manually
when using the `simple' makefile. It {\it may} be possible to update it via

\begin{verbatim}
  ./scripts/mkdeps > deps.mk
\end{verbatim}

from within the source directory.

The file \File{src/gkw.mk} is the main GNU makefile which should not
need to be modified in general, unless new variables are being introduced
to the code. Otherwise, all settings should be made via the files in
\File{config/}.

\subsection{Potential compilation issues}

Most compilers tested to date have no problems compiling the code.
Occasionally, some non-standard features or bugs get into the trunk version, creating
problems. However, these tend to get removed rapidly after they are discovered.
The Intel and GNU compilers are in most frequent use and are least likely to find unexpected new
issues when compiling and running.   On Hector, the PGI compiler has historically caused less problems than the Cray compiler.

Some compilers, such as those found on Fujitsu machines, may not like very
large source files; if such a compiler is used, it may be necessary to edit
the problematic source file and remove code which is not needed for the
particular run. %It is hoped that certain large source code files will be
%significantly broken up in the near future, thus preventing this issue.

Some compilers (e.g. Intel) do not by default include the present working directory
in the include path.  If \code{make} gives an error that \File{gkw_info.h} cannot be found,
when it already exists in the \File{/obj} directory, you should first try adding \code{I./} to 
\code{FFLAGS_INC} to your \File{config} file.

% ----------------------------------------------------------------------------
% ----------------------------------------------------------------------------

\subsection{\label{sec:additional-libraries}Additional libraries}

Here we list some third party libraries that can be used to extend the
functionality of the code.

\subsubsection{MPI-2}

An MPI-2 implementation is required for any distributed memory
parallel run.\footnote{The code might compile and run (with
  limited MPI-I/O and nonspectral functionality) with MPI-1 but this
  has not been attempted for a long time, and it is likely that some
  MPI-2 function calls would need to be commented to make this
  possible.}  MPI is enabled through preprocessing by setting
\code{MPI=mpi2} in the \File{config} file.  For most MPI
implementations, the include files and linking are automatically
handled by a compiler wrapper script such as \code{mpif90} or
\code{mpiifort}.

% ----------------------------------------------------------------------------

\subsubsection{FFTW3}

For nonlinear runs a Fourier transform library is
needed. At the moment only the FFTW3 library is supported.
If this library is installed it can be linked against by setting
\code{FFTLIB = FFT_FFTW3} in the corresponding \File{config} file of gkw.
Depending on the system you may also need to set \code{FFLAGS_INC} and \code{LDFLAGS}
to link to \code{-lfftw3 (-lfftw3f)} for double (single) precision.

% ----------------------------------------------------------------------------

\subsubsection{AMD and UMFPACK}
The two libraries \File{amd} and \File{umfpack} are needed for the (experimental) 
implicit time integration scheme and the nonspectral and global cases.
They are provided with the GKW trunk and must be compiled before GKW.

At present these libraries can only be compiled in double precision. 
If GKW is compiled in single precision, then type conversions are made 
wherever these libraries are used in the code.  
To compile and link these libraries to GKW one must add to the GKW config file
\begin{verbatim}
IMPLICIT = umfpack      # Compile GKW with umfpack (64 bit) or umfpack32 (32 bit)
\end{verbatim}
In most cases this should be sufficient; the libraries should then be built and linked automatically.   
One can also explicitly require compilation and linking of umfpack by setting
\begin{verbatim}
UMFPACK = compile       # Compile the UMFpack provided with GKW
LD_UMFPACK = -lumfpack  # link GKW with UMFpack
\end{verbatim}

% ----------------------------------------------------------------------------

\subsubsection{\label{subsubsec:slepc}SLEPc and PETSc}

For the eigenvalue solver the two libraries {\sc slep}c and {\sc pets}c are required
(where the former depends on the latter).
Note that these two libraries have to be compiled for using complex numbers
to work. This can be achieved with the option
\code{--with-scalar-type=complex} for the configure script of {\sc pets}c. {\sc slep}c
inherits these settings automatically from {\sc pets}c (if configured afterwards).
To link these libraries, one must add the paths for the libraries to the \code{LDFLAGS} together
with some subset of these (-lslepc -lpetsc -llapack -lblas -lX11 -lmkl) and
you must set \code{SLEPC = HAVE_SLEPC} in your \File{config} file for GKW.  
If the headers are not installed in a standard location for Fortran
you may also have to add the library include paths to \code{FFLAGS_INC}.  
It might also be necessary to add the path to petscconf.h to \code{FFLAGS_INC} 
(probably under an \File{include} subdirectory for each architecture).

% ----------------------------------------------------------------------------

\subsubsection{HDF5}\label{sec:hdf5}

The HDF5 library is needed to write output in the HDF5 data format.
Note that at the moment GKW expects the HDF5 library to be compiled in
its \emph{serial} version.\\
To link this library, one must add the paths ot the \code{LDFLAGS} and
\code{FFLAGS_INC} variables. It is a good idea to make use of the
\code{h5fc} wrapper which is provided by the HDF5 library to make
compilation easier (e.g. see the helios config).  Furthermore,
set \code{IO_LIB = HAVE_HDF5} in your \File{config} file for GKW, this
will then enable GKW's HDF5 support.

Note that the default output format of GKW depends on the availability
of HDF5. If GKW was compiled without HDF5 it then \code{io_format='mixed'} is the default, but if
compiled with HDF5 the default becomes \code{io_format = 'hdf5+ascii'},
see also section \ref{sec.diagnos-data-format}.

Note also that some datasets, in particular higher-dimensional (3D+)
diagnostic data (using MPI-IO) is not output in HDF5 format at the
moment, because parallel HDF5 is not yet supported by GKW.  If such
data is requested, it will be output in fortran binary output instead of HDF5.

% ----------------------------------------------------------------------------
% ----------------------------------------------------------------------------

\section{Running GKW}

Before you do any run it is important to realize that it is not all that hard to produce nonsense. 
The code is not under all conditions stable, and it is not always immediately obvious if a 
result is wrong. It is therefore important to diagnose your runs to make sure they are O.K.. 
It is your responsibility as a user to produce physically meaningful results.

Here it is assumed you have the necessary MPI environment set up and
running on the machine (consult your MPI library documentation) and that any run time
dependencies can be satisfied.

In order to run GKW, the code requires a single input file \File{input.dat} to be present
in the run directory. Optionally, a file containing equilibrium geometry information from
the CHEASE code can be provided as input.
To get started, there are a few example input files found in the \doc{input} directory,
together with corresponding \File{README} files. The file \doc{input.dat.sample} lists every input switch
together with a description of its purpose. \doc{input.dat.minimum} is a bare minimum input file containing
the essential inputs required for the code to run.

To run the code in parallel, you will probably need to use the program provided by the
MPI implementation; this is usually called \name{mpirun} or \name{mpiexec}.

GKW will try to select an appropriate parallelisation given the gridsize input but this can be overridden 
(by setting \name{N_procs_[sp|mu|vpar|s]} in the gridsize namelist).  Nevertheless it is
sensible to have in mind a parallelisation setup when you set the gridsize, as the user controls the number of 
processors the code will be run on and should choose a number that fits with the
gridsize.  For the largest parallel runs, the user is recommended to manually choose the parallel setup.

The rest of this section describes a basic input file.
Many of the input switches are optional and default values will be used if they are not provided.
The input variables are provided in Fortran namelists, for example the namelist `control' contains
the main switches for the code. The namelist examples presented here provide you with a basic set 
needed to run the code.  The namelists are sensitive to bad formatting, in which case the code should 
inform you which namelist has the problem.

GKW tries to catch as many invalid input combinations as possible. Most of the time if you input a set of 
incompatible switches the code will tell you and abort if necessary. For more minor errors the code will generate 
a warning that it is ignoring an input parameter and continue.  The actual input parameters the code runs with are 
written to \File{input.out}.  By renaming \File{input.out} to \File{input.dat} the code should re-run without the warnings
(if the compiler correctly formats namelist outputs).

\subsection{Modes}

The spacing and location of the wavevectors is determined by the `mode' namelist.
For \name{mode_box=.false.} there is only one radial wavevector considered, but the field line can be extended to 
take more than one poloidal turn by setting \name{nperiod}. The total number of full poloidal turns is then 2 \name{nperiod} - 1. 
The bi-normal modes are determined by the input \name{kthrho} which specifies the value(s) of
$k_\theta\rho_{\rm ref}=\sqrt{g^{\zeta\zeta}(s=0)}k_\zeta$, see Eq.~(\ref{kperp}). For {mode_box=.false.} the bi-normal modes are given as a list i.e.
 \name{kthrho=0.1,0.2,0.3} (the variable \name{krhomax} is ignored).
The radial mode is determined by the input \name{krrho} which specifies the value of $k_r\rho_\textrm{ref}=\sqrt{g^{\psi\psi}(s=0)}k_\psi$.
Alternatively, a poloidal shift of the balooning angle can be specified using the input \name{CHIN}. A finite value of $k_\psi$ will then be
calculated to minimise $k_\perp\rho_{\rm ref}$ at the poloidal angle $\theta = \name{CHIN}$. The value of $k_\psi$ is given by
$$k_\psi = - k_\zeta g^{\psi\zeta}/g^{\psi\psi}$$
where the metric elements are evaluated at the poloidal location specified by CHIN (poloidal angle, from the magnetic axis, positive counterclockwise, in radian and
zero at the low field side midplane) and $k_\zeta=k_\theta\rho_{\rm ref} / \sqrt{g^{\zeta\zeta}(s=0)}$, as usual.

For \name{mode_box=.true.}, the field line must have one poloidal turn, i.e. one must set \name{nperiod=1}. In this case 
the radial modes are coupled at the end of the field line according to the ballooning shift connected with the
magnetic shear.  
The bi-normal modes will be equally spaced from $k_\theta\rho_{\rm ref} =0$ to \name{krhomax} (the \name{kthrho} input is ignored).  
The integer $i_k$ in Eq.~(\ref{ikxspace}) is input into the code as \name{ikxspace} and determines the balance between radial 
resolution and radial box size.  \name{ikspace} should be adjusted to give an approximately square box (the code will output LX and LY).
\begin{table}[h!]
\begin{center}
\begin{tabular}{c|c|c|c|c}
 NS & ikxspace & NX & NPERIOD & mode_box \\
\hline
 80      &    N/A   &  1  &    3    & false    \\
 16      &     1    &  5  &    1    & true     \\ 
\hline
 $n_s.(2n_p-1)$ &    N/A   &  1  &    $n_p$    & false    \\
 $n_s$    &     1    & $2n_p-1$ &  1   & true   \\

\end{tabular}
\caption{Equivalent formulations example with \name{mode_box} true and false}
\end{center}
\end{table}\\

\subsection{Dissipation \label{sec.dissip}}

The parameter \name{disp_vp} sets the dissipation on the trapping term. 
It acts a bit like a collisionality but the normalization is different: The dissipation decreases at fixed \name{disp_vp}
when \name{N_vpar_grid} is increased, so that convergence should always be reached.  For all collisionless problems
that depend critically on the trapping (TEM modes, particle fluxes) one should not set the dissipation in velocity space too high.
For these cases a lower \name{disp_vp} (say 0.01) will allow a lower resolution convergence of the particle fluxes.  
When collisions are included, \name{disp_vp} can and should be small (0.2).

The parameter \name{disp_par} sets the dissipation on the streaming term, and 0.2 is a reasonable value for most linear runs.
For nonlinear runs, values in the range 0.5 - 2.0 are typically required for numerical stability, 
although stable results can be achieved with lower dissipation by using the Arakawa scheme.
If parallel dynamics must be finely resolved, higher \name{N_s_grid} bay also allow lower \name{disp_par}.
The default form of the parallel dissipation uses \name{idisp = 2}, which does not vary over the velocity grid
and is usually more stable for electromagnetic runs .  However, 
in some physical cases with a sharp parallel structure (usually at low $k_y$),
excessive dissipation appears to drive a linear numerical instability;
in these cases the stability may be improved by using  \name{idisp = 1}. (See \issue{201}).

The circular geometry tends to be more stable than the $s-\alpha$ geometry if numerical problems arise in the zonal mode \cite{CAS10}.
For runs with strong $E \times B$ shear, \name{disp_par} should be minimised, and some radial dissipation (\name{disp_x}=1.0) should 
be used (see Ref. \cite{Casson-Thesis} or \cite{exb-dissip}).  \name{disp_x > 0.0} is also required for the finite differences of the
nonspectral version.  \name{disp_y > 0.0} can be useful in some nonlinear runs to cut out high k physical instabilities (e.g ETG modes) and 
supress high-k spectral pileup.

\subsection{Gridsizes}

A detailed GKW linear convergence study including many figures is available in Ref. \cite{VET10}, of which the numbers below are a summary.
The gridsizes required for a converged result will vary for each problem, but here we give some guidance assuming \name{disp_vp}=0.01 and circular geometry.  For adiabatic runs, $[N_s,N_\mu,N_{v_\parallel}]=[16,8,16]$ is often sufficient (Here, $N_s$ is points \textbf{per poloidal turn}, but \name{N_s_grid} in the input file is the total number.). 
Collisionless ITG modes with kinetic electrons will give growth rate and frequency converged to within 5\% for $[N_s,N_\mu,N_{v_\parallel}]=[16,8,32]$, 
but if the particle fluxes are to be examined, gridsizes of $[N_s,N_\mu,N_{v_\parallel}]=[40+,8,48]$ 
may be required (this is because the trapped passing boundary must be accurately resolved).

In general, trapped electron modes require more resolution to get accurate growth rate and frequencies, 
but once these are resolved, the particle fluxes should be also.  A TEM mode is converged to with 5\%
for $[N_s,N_\mu,N_{v_\parallel}]=[30,8,64]$  Including collisions can reduce the values of $N_{v_\parallel}$ required 
substantially ($N_{v_\parallel} = 16-32$), with the lower figures for greater collisionality.

For a linear run, \name{Nperiod} will depend on the type of mode you are looking at.  For some ITG cases with high magnetic shear, 
3 would be sufficient, but for some TEM cases or cases with lower magnetic shear, as many as 10 may be required.  Remember to adjust 
\name{N_s_grid} accordingly.

For a nonlinear run, \name{Nmod} should be at least 16.  In the past a
value of \name{nmod}=21 (or 43, 86,...) was used for particularly
efficient FFT computations, but this is no longer significant.  For
kinetic runs \name{NX}=167 is a sensible size.  For adiabatic runs
\name{NX}=83 may be sufficient.  For cases with very low magnetic
shear or high E x B shear, \name{NX}=339 is sometimes used for
convergence tests.

For $N_s$, more points are required if using complex general geometry.  
The boundaries of the velocity space do not usually need to be increased from the defaults.

\subsection{Species}
There is no default species data, hence the user must provide some. The species input must satisfy quasi-neutrality. This requires the user to set ratios of densities and charge to satisfy
\begin{equation}
 \sum_s Z_s n_s =0, \qquad \sum_s Z_s n_s {1 \over L_{n_s}} = 0
\end{equation}
Multiple ion species can be input but only one negatively charged species is allowed.  The number of species read in is controlled by the number of species parameter in the `gridsize' namelist and
hence it is possible that later species data you specify will not be read in if this is set incorrectly.  The \name{number_of_species} input variable does not include adiabatic electrons (the
species data for the negative charge species will still be read).

Note that for a trace species, $n_s=0$ is acceptable. The density only appears in the field equations and for scattering species in collisions.  A trace particle does not contribute to the fields and does not act as a scattering species, so setting $n_s=0$ will give the desired result.

\subsection{Stability}
The timestep required for numerical stability can vary greatly depending on input parameters, and for kinetic electrons will need to be much smaller.
For a linear run, there is a time step estimate made at the start of the run and the code will post a warning if
the chosen timestep is larger than this. For a non linear run, the minimum time step is estimated based on the
non linear terms, then the time step is adapted as necessary. A minimum time step size can be set in the code (\name{dt_min}) so
that the run stops when the time step becomes too small. The parallel structure of the solution is a good
indicator of a the stability of a solution.

For linear runs a stable solution will converge to a constant growth rate. A tolerance for the growth rate can be
specified in the code input (parameter \name{gamatol}), so that the run ends when the standard deviation of the growth rate computed over the last six big time steps is less than \name{gamatol}. When the tolerance on the growth rate is used to end the run, it is advised to keep the big time steps long enough to have a meaningful standard deviation. Typically, if the big time step length is such that $\name{DTIM}\times\name{NAVERAGE}>0.2$, setting $\name{gamatol}=10^{-6}$ will usually ensure convergence.

GKW outputs the global growth rate
in \File{time.dat} so a growth rate spectrum can be generated from a batch of linear runs with 
\name{NMOD=1} and \name{mode_box=.false.}.  Typical grid sizes can be seen in the example input files.

\subsection{\texorpdfstring{$\beta$ and $\beta'$}{beta and beta prime}}
\label{Sec:betabetapr}
The plasma $\beta$ enters the field equations (parallel and perpendicular components of the Ampere's law, Eq.~\ref{eq:AmpPar} and
\ref{eq:AmpPerp}) through the parameter $\beta_{ref}$ defined in Eq.~\ref{eq:betaref} and recalled here:
\begin{equation}
 \beta_{ref} =  \frac{n\ind{ref}T\ind{ref}}{B^2\ind{ref}/2\mu_0}.
\end{equation}
The parameter $\beta_{ref}$ is linked to the total plasma $\beta$ at the reference magnetic field as follows:
\begin{align}
\label{eq:betadef}
 \beta_0 &= \sum\ind{species} 2\mu_0\frac{n_sT_s}{B^2\ind{ref}} = \beta_{ref} \sum\ind{species} \frac{n_s}{n\ind{ref}}\frac{T_s}{T\ind{ref}} = \beta_{ref}
\sum\ind{species} n_{N_s} T_{N_s}\\
& = \beta_{ref} \sum\ind{species} \beta_{N,s}\qquad\text{with}\quad \beta_{N,s} := n_{N_s} T_{N_s}
\end{align}
The user can specify the value of $\beta_{ref}$ used by GKW within the namelist \name{SPCGENERAL}. Two options are implemented and controled by
the switch \name{BETA_TYPE}:
\begin{itemize}
 \item \name{BETA_TYPE=`ref'}\\
  The value of $\beta_{ref}$ is given as an input in \name{BETA_REF}.
 \item \name{BETA_TYPE=`eq'}.\\
 The total plasma $\beta$ at the reference magnetic field is read from the MHD equilibrium file and $\beta_{ref}$ is calculated using Eq.~\ref{eq:betadef}
and the species parameters.  The value given as an input in \name{BETA_REF} is not used and replaced by the calculated value. Note that this option is
only available if \name{GEOM_TYPE='chease'} in namelist \name{GEOM}. If it is not the case a warning is displayed and $\beta_{ref}=0$ is used instead. 
\end{itemize}

In addition, the radial derivative of the plasma $\beta$ modifies the curvature drift, as seen in Eq.~\ref{eq:drifts}. Here again, the value at the
reference magnetic field is needed (the dependence on the magnetic field is introduced directly in \File{linear_terms.F90}): 
\begin{equation}
\label{eq:betapr0def}
 \beta'_0 = \frac{2\mu_0}{B\ind{ref}^2}\sum\ind{species}\frac{\partial p_s}{\partial \psi} = - \beta_{ref} \sum\ind{species}  n_{N_s} T_{N_s}
[R/L_{n_s} + R/L_{T_s}]
\end{equation}
Three ways of inputing $\beta'_0$ are implemented:
\begin{itemize}
 \item \name{BETAPRIME_TYPE=`ref'}\\
The value of $\beta'_0$ is given as an input in \name{BETAPRIME_REF}.
 \item \name{BETAPRIME_TYPE=`sp'}\\
The value of $\beta'_0$ is made consistent with the species parameters and the value of $\beta_{ref}$ defined above
using Eq.~\ref{eq:betapr0def}. The value given as an input in \name{BETAPRIME_REF} is not used and replaced by the calculated value
 \item \name{BETAPRIME_TYPE=`eq'}. \\
The value of $\beta'_0$ is read from the MHD equilibrium file. The value given as an input in \name{BETAPRIME_REF}
is not used and replaced by the calculated value.
\end{itemize}

Note that when the species parameters are used, ALL species are considered including the adiabatic one.

Reminder: if \name{GEOM_TYPE='chease'}, the values of $R\ind{ref}$ and $B\ind{ref}$ used in GKW are the values \name{R0EXP}
and \name{B0EXP} given as an input to CHEASE. It is advised that \name{R0EXP} and \name{B0EXP} are chosen to be the radius of the magnetic axis and
the value of the magnetic field at this position respectively for better convergence in CHEASE.

\subsection{Generating a linear growth rate spectrum}
GKW currently outputs the global growth rate (and frequency) in \File{time.dat} so the easiest way to generate a growth rate spectrum is do a batch of linear runs with \name{NMOD=1}, \name{mode_box=.false.}, and \name{NX=1}, and scan over \name{kthrho} and take the growth rate from each. 
% If \name{mode_box=.true.} and you have mode than one radial mode, 
% scan over \name{krhomax} to set the toroidal wavevectors (to keep the kx spacing the same you also have to vary ikxspace with krhomax); 
The matlab scripts \matlab{create_gkwscan.m} and \matlab{growth_rate_spec.m} help with creating the input and plotting the output; 
script \scripts{gkwnlin} runs the batch jobs.

Alternatively, a growth rate (frequency) spectrum can be generated by running with \name{mode_box=.true.} and setting
\name{lgrowth_rates=.true.} (\name{lfrequencies=.true.})in the \name{DIAGNOSTIC} namelist.  For each toroidal mode,
the growth rate and of the mode with $k_\psi=0$ will be output.  To get this spectrum to exactly match a spectrum
generated by the first method at higher $k_\zeta$ will require large \name{NX} (and low \name{ikxspace}) to fully resolve 
the mode along the field line.  The excercise of matching the spectra generated by these two methods
will require the user to understand the mode coupling over the parallel boundary condition
and will indicate the \name{nx} resolution required for \name{mode_box=true} and nonlinear runs.

\pagebreak

\subsection{Restarts \label{sec.restart} and checkpoints}

On a successful exit, GKW writes a binary file \File{FDS} which
contains the distribution function needed for a restart.  Along with the
input file, this is the only file required to restart the code.  
For seamless continuation, some other quantities such as file counters 
and normalized time are also written to the ASCII file \File{FDS.dat} which
contains Fortran namelist which the code reads.   The binary restart file 
is written using MPI-IO and is independent of processor layout.  This means that
runs can be restarted with different parallelizations from the original run.

GKW will only attempt to restart from the above files if \name{read_file=T} in the \name{CONTROL}
namelist.  If it does not find any files, it will initialize a new run.  Under the default behaviour,
when this subsequent run completes, a restart files \File{FDS} and \File{FDS.dat} will be overwritten.
This behavior can be altered by setting the integer \name{irun} in the \name{CONTROL} namelist which
allows incremental numbering of restart files.  If \name{irun=1},  
\File{FDS} is read if present and \File{FD1} is written.   If \name{irun=2},  
\File{FD1} is read and \File{FD2} is written, and so on.

For systems on which MPI-IO is not available or problematic\footnote{
For large problems with more than about 64 processes,
there can be problems/errors when reading/writing the restart files in
parallel, depending the filesystem and the MPI implementation.
This has been observed on some Cray XT4 machines, where in
some cases it was necessary to modify environment variables to
successfully write restart files. There is also a workaround in the
code which must be set at compile time via adjustments to two variables in
\File{switches.F90}: Rather than writing to file with all processes
simultaneously, this allows writes with subsets of processes
at a time, where the file is closed then re-opened between subsets.
Increasing the number of subsets reduces overheads and can allow files to
be written successfully. This might also improve the I/O performance on
machines, where such operations already work. 
However, if the number of subsets is too large, the I/O
performance will suffer. The variable \name{min_fdisi_io_subsets} selects
the minimum number of process subsets to use. When problems are encountered, 
it is suggested to first try setting this value to the value of \name{n_procs_sp}, 
then to \name{n_procs_sp*n_procs_s} if problems persist. Using larger values will
likely give very poor performance, but may be necessary. The other variable,
\name{min_procs_io_split}, sets the minimum number of processes for which
the read/write split should be used.},
GKW can also be made to write
a legacy restart file format, in which each processor writes its own file 
\File{FDSXXXXXX} numbered by processor (option \name{restart_file_version=1} in \name{CONTROL}).  
In this case, restarts may not alter the parallelization.

Intermediate checkpoints can also be written by setting the integer \name{n_dump_ts=50} in the \name{CONTROL} namelist,
to write, for instance a checkpoint every 50 large timesteps.  This is strongly advised for any expensive nonlinear simulation.  
Two restart files \File{DM1/2} are alternately overwritten by a newer one each time.  These files have exactly the 
same format as the final restart file \File{FDS} and can be used for a restart by simple renaming.
One should note that if restarting an interrupted run from a checkpoint,  some the time trace files may have duplicated lines, 
which can be easily verified by checking or plotting column 1 of \File{time.dat}.

A clean exit can be requested at anytime while the code is running by creating a
file named \name{gkw.stop} in the run directory. An additional checkpoint 
file can be created by instead creating a file named \name{gkw.dump}, after which
the code will continue to run.

Some batch systems will automatically resubmit a job which is terminated by a hardware problem; 
to allow GKW to automatically restart from the \File{DM1/2} files (and take precedence over FDS files) if present, 
set \name{auto_restart=T} in the \name{CONTROL} namelist.   GKW could also be made to interface with batch 
systems which are provide accounting flexibility for codes able to halt on request,
by means of the \name{gkw.stop} file creation mechanism.

\pagebreak

\subsection{Tips and tricks}

%Should this be a wiki page?

This section aims to document other useful things to know when running the
code. Please add here anything else useful to the user that we have yet to
document properly. Received wisdom, etc.

\begin{itemize}

\item The linear timestep estimator is based on the Courant condition for the
       Vlasov equation, and the frequency of shear-Alfv\`{e}n wave \cite{LEE01}, so it is not
       foolproof for some high frequency waves (e.g. ETG). If GKW gives unphysicaly large growth
       rates, the first thing one should try is to reduce the timestep (no more
       than a factor of two at a time). If your problem requires a very small
       timestep (less than 1e-6), something else is likely to be wrong, but if
       a very small timestep is really required it may be best to run with
       \name{normalized=.false}. Equally, for most efficient resource usage,
       one should try to find the upper limit of the stable timestep.

\item For nonlinear kinetic runs, adding a small electromagnetic component e.g.
       $\beta=0.03\%$ can stabilise some long wavelength instabilities
       (electrostatic shear-Alfv\`{e}n waves), allowing a much larger stable
       timestep, as suggested in \cite{MER10}. 
       %Same trick is mentioned in Jenko 2010 paper

\item For kinetic electron runs (\name{adiabatic\_electrons=.false.}) the default 
      initialization option \name{finit='cosine2'} may lead to 
      initial bursts. A different initialization of the distribution function, 
      e.~g., \name{finit='cosine5'}, may then be more appropriate.

\item The nonlinear timestep estimator rarely manages to prevent blowup in the
       event of numerical instability, but can be useful for strongly nonlinear
       physics.   For very large numbers of processors it may cause an overhead
       which reduces the scaling efficiency.  It can therefore be disabled
       by setting. \name{nl_dtim_est=F} in the \name{CONTROL} namelist, which
       will cause the code will abort rather than continue if the timestep
       estimate is violated.  The code could then be restarted from the previous
       dumpfile with the nonlinear timestep estimator enabled, or with a smaller
       timestep.

\item The nonlinear timestep estimator is often too conservative.  
      If your problem is limited by the nonlinear timestep, 
      it may be possible to save CPU resources by increasing it (no more than $90\%$) 
      by setting e.g. \name{fac_nl_est = 1.6} in \name{CONTROL}.

\item Don't be tempted to use more cores than you need.  Pushing the parallelisation to its limit 
       (particularly less than 4 points per processor in parallel velocity) will result in inefficient scaling.

\item Be aware of the effects of the boundary condition at the end of the field line. If not running a box of modes, you should
       check that \name{nperiod} is large enough so that the results are insensitive to the free boundary. If running with
       \name{mode_box}, care should be taken with the radial mode resolution so that most toroidal modes have several periods and
       do not feel the effect of the boundary. This is particularly important when kinetic electrons are used.

\end{itemize}

\subsection{Hybrid scheme \label{sec:hybrid}}

Hybrid parallelism combining distributed memory MPI parallelism with shared memory OpenMP parallelism
has been implemented to allow the code to scale further, and more efficiently. 
Many modern high performance machines now have an architecture with a number of shared memory CPU cores on each node (2-24), 
and a large number of these nodes connected by a fast network interconnect.  
The communication between nodes is much slower than between cores, and a well optimised and well
utilised MPI implementation will take advantage of this by putting the most intensive communication on the local node.
  In cases where this is not done, or cases where a subset of 
the problem grid does not fit nicely on a single node, the efficiency of the inter-node communication may be increased by reducing 
the number of MPI processes involved in the intra-node communication.  For example, on a machine with 8 cores per node (e.g. HPC-FF), 
1024  processors can be utilised with the combinations 
(number of MPI processes, number of OpenMP threads per node) = (1024,1), (512,2), (256,4), or (128,8), 
where the lower number of MPI processes in the latter case may result in a more efficient reduction sum 
(though in theory the amount of inter-node data communicated by the ideal MPI implementation would remain the same).

In GKW, hybrid parallelism (for the spectral scheme only) is achieved by subdividing the most CPU intensive loops amongst OpenMP threads.  
The slowest of these is the loop in nonlinear terms, where for each local grid point, 5-7 FFTS in 2D must be performed. 
At the maximum limit on the number of MPI processes, 
there will always be at least 2 (4 with collisions) points in this loop (it contains \name{NMU*NSP*NS} local grid points) 
which can be divided between threads.  
(The code could be rewritten here to increase the number of threads that can be used in this section, 
if the hybrid scheme is demonstrated to bring a clear benefit at these numbers of threads).  
The second largest loop which must be parallelised is the sparse matrix-vector multiply involved in the linear terms. 
Due to the sparse matrix storage format, this loop cannot be vectorised, but cache efficiency can be maximised 
by an appropriate sorting of the matrix.  After this sort, the matrix elements are subdivided between threads at carefully chosen boundaries,  
to ensure there is no data race in the update of the right hand side.  The same technique is used for the fields calculation. 
All possible smaller loops ($<$5\% runtime) are also parallelised with OpenMP, 
but here there are diminishing returns here due to the overhead involved in thread branching.  
This inherently serial element of the code will set the limit of the number of threads that can efficiently be used (likely four at present).

The hybrid scheme with OpenMP parallelism can allow GKW to scale better in certain situations for large runs.  
Tests on HPC-FF and Babel Bluegene machines (combined with some reasoning), suggest that the cases in which it is most likely to be useful 
are those in which the MPI communications are the bottleneck.  When this occurs depends on the problem and the hardware.  
These situations are likely to include one or more of the following:
\begin{itemize}
 \item Problems with large numbers of modes
 \item Problems with collisions (4 threads has been demonstrated to bring a benefit here)
 \item Runs on very large numbers of processsors, near the limit of GKW MPI capability
 \item Hardware with slow communication (HPC-FF or BlueGene as opposed to Hector)
\end{itemize}

Preliminary tests with the hybrid scheme on up to 1536 processors (on HPC-FF: IBM-Intel-Xeon-Infiniband) demonstrated up to a 5\% 
performance increase using two OpenMP threads per node.  More extensive scaling tests on Bluegene/P BABEL have
demonstrated a scaling efficiency increase of up to 20\% with 4 threads when running a collisional case on 8192 cores and above 
(for more details see Sec. \ref{sec:bluegene}).
%ALSO, RECREATE HECTOR 2a or new HECTOR2b or HECTOR 3 scaling test to see in open MP can bring a benefit there.
%Want to do a large collisional scaling test without the diagnostics on.
%Want to know how much is due to momentum conservation (this was not even integrated into hybrid shcheme at time of test).

You are advised to do some brief checks for your problem and hardware to determine
if the hybrid scheme is an advantage.  Best results will probably be obtained using the hybrid scheme 
with no more than four OpenMP threads per MPI process.
 For acceptable results, the number of threads must be a factor of the local processor gridsize 
 \name{NMU*NSP*NS} (which is also the maximum number of threads), as a loop of this size is parallelised in the nonlinear terms.

To use the hybrid scheme, you need a compiler that supports OpenMP.  You will need to compile with
SMP=OPENMP preprocessor flag, openmp compiler flag, and link with the correct libraries (by using the same compiler flag). 
In the standard makefile build process, this is usually achived by setting 
\begin{verbatim}
SMP=OPENMP
FFLAGS_OMP=-openmp (for intel, or -fopenmp for gnu)
\end{verbatim}
Using the compile optimisations specific to the processor is also recommended.

\subsection{A brief guide to moving to global simulations}
\label{sec:brief-guide-moving}

\emph{This guide needs improvement, please add to it.}

This section assumes you can already cope with running non_spectral.
Running global simulations is somewhat more complicated:

The input switches for the global version are not yet very well
documented, and the input checks and error messages may not yet be
very comprehensive or informative (there is a job there for anyone 
that wants to be helpful).

It is probably easier to start from an existing global test case, 
but to move from local to global you would need to add at least 
the following switches.

\paragraph{CONTROL}
\subparagraph{\name{spectral_radius = .false.}}
Since the radial direction is no longer 
homogeneous you can not use a spectral 
method 
\subparagraph{\name{flux_tube = .false.}}
This is the switch that tells the code 
it is in fact a global run 
\subparagraph{\name{disp_x = 0.05}}
You need this (or more). Since there 
are radial derivatives in the equation 
one needs to put some dissipation.

\paragraph{GRIDSIZE}
\subparagraph{\name{vpmax = 4.0}} The velocity grid does not vary with radius, 
and so it must be able to resolve both the low 
as well as the high temperatures in the radial 
domain. This means that one often has to increase 
both the maximum value of the grid (vpmax) as 
well as the velocity space resolution (n_vpar_grid)
The value chosen depends on how much the temperature 
varies over the radial domain.
\subparagraph{\name{n_vpar_grid = 64}} You must often choose this larger than a fluxtube 
run (see comment vpmax)
\subparagraph{\name{mumax = 8.0}} You must often make the perpendicular velocity 
grid larger (see comment vpmax) 
\subparagraph{\name{n_mu_grid = 16}} also the resolution in the perpendicular velocity
must be larger usually (see comment vpmax)
\subparagraph{\name{psil = 0.02}} inner boundary eps=r/Rref
\subparagraph{\name{psih = 0.3}} outer boundary eps=r/Rref
\subparagraph{\name{nx = 128}} Better to be even in nonspectral. Also this value 
must be chosen high enough such that you resolve 
at least 2 times the ion Larmor radius (1 is better
and kinetic electron runs might require more 
resolution (this is true for nonspectral as well) 
\subparagraph{\name{n_procs_x = 4}} Now you can parallelize here too

\paragraph{MODE}
\subparagraph{\name{n_spacing = 2}} For a run with one mode this is the toroidal mode 
number (kthrho / krhomax are not used). For a 
nonlinear run with many modes this is the mode 
spacing, i.e. n_spacing = 2 with nmod = 4 means 
that the toroidal modes 0, 2, 4, and 6 are kept 
in the simulation (this is the half-torus)

\paragraph{SPCGENERAL}
\subparagraph{\name{rhostar = 1e-3}} rho_ref/R_ref

\paragraph{SPECIES}
\subparagraph{\name{dens_prof_type = 'exp_tanh'}}
\subparagraph{\name{temp_prof_type = 'exp_tanh'}}
\subparagraph{\name{temp_prof_coef = 0.0,0.0,0.0,0.0,0.0}}
\subparagraph{\name{dens_prof_coef = 0.0,0.0,0.0,0.0,0.0}}
These define the radial profile of temperature/density, and through the
derviatives also the profiles of the density gradient lengths. For possible
profile types and the meaning of the coefficients for those, see first part of
section \ref{secprofilefunctions}.

\paragraph{GEOM}
\subparagraph{\name{prof_type = 'parabolic'}}
\subparagraph{\name{qprof_coef = 0.0,0.0,0.0,0.0,0.0}}
These define the radial profile of the safety factor, and through the
derivative also the shear profile. For possible profile types and the meaning
of the coefficients for those, see second part of section \ref{secprofilefunctions}.

The chosen profiles are output in prof_back.dat 
(and there is a matlab script to read them).

If you are going to attempt a nonlinear simulation, then you will
probably also want the krook operator to prevent profile relaxation.
See section \ref{seckrookoperator}, for more details.

\paragraph{KROOK}
\subparagraph{\name{krook_option=1}}, global krook preserving momentum
\subparagraph{\name{gammak=0.05}}, global krook damping rate (should be << linear gr)
\subparagraph{\name{nlkrook=.true.}}, global krook
for boundary krook cells, see input.dat.sample

You should also take a look at the \name{GYROAVERAGE} namelist, which
is fairly well documented in \doc{input.dat.sample}.

\pagebreak

\section{Example GKW input file}
\vskip 0.2 truecm
Further example input files are provided with the code, including a comprehensive input file 
\doc{input.dat.sample}, a minimal input \doc{input.dat.minimum}, 
and input files for some common benchmark cases in \doc{input/}.
\begin{scriptsize}
\begin{verbatim}
 &CONTROL                   !Namelist read in module CONTROL
  non_linear = .false.      !Include the non linear terms, computationally expensive fft
  zonal_adiabatic = .false.,!If zonal flows correction included for adiabatic electrons
  order_of_the_scheme = 'fourth_order'  !or 'second_order' in space
  dtim   = 0.02          !Time step size (normalised time)
  ntime  = 25,           !Number of iterations of NAVERAGE.
  naverage = 100,        !Number of timesteps between re-normalisation, data output
  nlapar = .false.,      !true = keep parallel electro-magnetic potential.(electrostatic=false)
  nlbpar = .false.,      !true = keep compressional magnetic field (electrostatic=false) 
  collisions = .false.   !Turn on the collision operator
  disp_par = 1.0         !Dissipation coefficient for parallel derivatives.
  disp_vp  = 0.0         !Dissipation coefficient for parallel velocity space
  max_seconds = 1000.    !Internal program kill time.
  gamatol     = 0.       !tolerance in the linear growth rate for which the code should stop
  io_format   = 'ascii'  !chose output format to be ascii
 /
 &GRIDSIZE        !Namelist read in module GRID
  nx = 83,        !Number of radial wave vectors - needs to be an odd number for fft
  n_s_grid = 50,  !Number of grid points along the field line 
  nperiod = 1,    !Integer - the field line makes 2*nperiod - 1 poloidal turns. 
  n_mu_grid   = 8,        !Number of magnetic moment grid points
  n_vpar_grid = 32,       !Number of grid points for parallel velocity 
  nmod = 21,              !Number of bi-normal modes (k_zeta)
  number_of_species = 1   !Number of kinetic species. 
 /
 &MODE                !Namelist read in module MODE
  kthrho  = 0.5,1.0,  !'Poloidal' wavevector(s) if mode_box=.false.
  mode_box = .true.,  !Determines if there is a 2D grid of modes. If true use nperiod = 1
                      !Also determines if radial modes are coupled at end of field line.
  krhomax = 1.0,      !For mode_box, the maximum 'poloidal' wavevector (k_perp rho_i) used
                      !k_perp evaluated on the low field side of the outboard midplane
                      !rho_i evaluated on the flux surface at the major radius of magnetic axis
  ikxspace = 7,       !Spacing between radial modes (resolution against box size)
 /
 &SPCGENERAL                   !Namelist read in module COMPONENTS
  beta_ref = 0.000,            !Plasma beta (not used if both nlapar, nlbpar are .false.)
  adiabatic_electrons = .true. !Kinetic electrons require a smaller timestep
 /
 &SPECIES       !Namelist read in module COMPONENTS
  mass  = 1.0,  !Species mass in terms of reference mass
  z     = 1.0,  !Species charge. If negative, assumed to be the electrons.
  temp  = 1.,   !Temperature of species scaled by reference temperature 
  dens  = 1.,   !Density of species scaled by reference density 
  rlt   = 8,    !Temperature gradient R/LT
  rln   = 3.5,  !Density gradient R/Ln
  uprim = 0.0,  !Gradient of the toroidal rotation Rref^2 grad Omega / v_thref
 / 
 &SPECIES 
  mass  = 2.72e-4, !Electrons
  z     = -1.0,    !Only one negatively charged species is allowed.
  temp  = 1.0, 
  dens  = 1.0,
  rlt   = 7.9,
  rln   = 3.5, 
  uprim = 0.0,
 /
 &GEOM          !Namelist read in module GEOM, s-alpha is default geometry
  shat = 1.0,   !Magnetic shear
  q    = 2.0,   !Safety factor
  eps  = 0.1    !Radial Coordinate
 /
 &ROTATION                   !Namelist read in module ROTATION
  vcor = 0.0                 !Rotation of the plasma vcor = Rref Omega / vthref = \Omega_N
  shear_rate = 0.0           !Normalised shearing rate for the E x B perpendicular shear 
  shear_profile  = 'none'    !To include a shear flow, use `wavevector_remap'
  cf_trap = .true.           !Include the centrifugal trapping effects           
  cf_drift = .true.          !Include the centrifugal drift
 /
 &COLLISIONS    !Namelist read in module COLLISONOP if collisions = .true.
  rref = 1  !reference major radius used in collision operator 
  tref = 1  !reference temperature in units of kev used for the collision operator
  nref = 1  !reference density in units 10^19 m^-3 used for the collision operator
  mom_conservation = .false.     !Use the correction to conserve momentum
  mass_conserve = .true.,        !Forces zero flux boundary on velocity space so no out flow of mass   
  freq_override=.false.          !Manually specify ii collision frequency to override ref values 
  coll_freq=6.515E-5             !Ion-ion collision frequency used if freq_override=.true.
 /
\end{verbatim} 
\end{scriptsize}

% related to auctex mode and latex-preview-mode in Emacs:
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "doc"
%%% End: