### 2022-11-05 bachelorthesis/01-introduction.tex
#39:  New structure of files and references

% 1. Introduction

\chapter{Introduction}
\label{chap:intro}