### 2022-11-05 bachelorthesis/02-theory.tex

% 4. Theory

\chapter{Theory}
\label{chap:theory}

% Text

% etc.