### 2022-11-05 bachelorthesis/03-methods.tex

% 3. Methods and MAterial

\chapter{Methods and Material}
\label{chap:methods}