### 2022-11-05 bachelorthesis/04-results.tex

% 4. Results

\chapter{Results}
\label{chap:results}

% Text

% etc.