### 2023-04-29 bachelorthesis/04-declaration.tex

\chapter*{Declaration}
\label{chap:declaration}

The author declares that the material presented in this thesis is his own work, unless explicitly stated otherwise. Chapter \ref{sec:convergence} and \ref{sec:threshold} is based on
\bigskip
\begin{itemize}
    \item [] \bibentry{Lippert2023}
\end{itemize}
\bigskip
and is an further iteration of the brief communication published in the journal \enquote{Physics of Plasma}. It provides additional plots and paragraphs that were not necessary to be included in the final publication. The format was changed as well because the publication was written in REV\TeX~4.1 and this bachelor thesis in \texttt{scrreprt}. The brief communication can be found in Appendix \ref{append:paper}.