### 2023-05-01 bachelorthesis/20-theory.tex
#39: New title style; corrected Chapter names

% 2. Theory

\chapter{Plasma Physics, Zonal Flows and Gyrokinetic Theory}
\label{chap:theory}

\thispagestyle{empty}
\newpage