### 2023-05-01 bachelorthesis/22-confinement.tex

\newpage
\section{Magnetic Confinement in Tokamak}
\label{sec:confinement}

In a tokamak devices strong magnetic fields confine the hot plasma. As mentioned in Chapter \ref{sec:motion} a magnetic field forces a perpendicular particle motion and a motion containing gyro motion and slow perpendicular gyro center drifts. Because of the much smaller size of the Larmor radius compared to the device size $R_0$ the particle and energy losses are caused by the gyro center drift. To avoid this at the end of the field lines the magnetic field in the tokamak devices is shaped like a torus. This type of geometry has nested surfaces with constant magnetic flux, so-called \textit{flux sufaces} and with magnetic field lines which appearing on that surfaces.