### 2023-05-18 briefcommunicaton/responseletter.txt
#37: Update Graphic, Refs + Responseletter

Editorial Office Comments:

* Revised manuscript source files must be in Word or LaTeX/ReVTeX. Replace the current PDF file with an acceptable manuscript source file.

    > [ ] Remove graphics hack (?)

* Separate figure files (separate from text and tables) are required for production purposes. Submit a separate figure file for each figure cited in the manuscript, not just one file containing all figures. The allowed file formats for figure files are TIF, PS, EPS, JPEG, or high-quality PDF. Figures may remain embedded in the manuscript. Number your separate figure files as they appear in the manuscript. For example 1, 2, 3. 
NOTE: DOC or DOCX figure files are not permitted.

    > [ ] Rename figure files

Reviewer Comments:

Referee #1:

This manuscript presents a dedicated convergence study of zonal flow generation from ITG turbulence in the Cyclone Base Case. It is clearly written and well-organized. Although it doesn't contain any new physics by itself (ExB staircase in a local gyrokinetic simulation is nothing but a stylish expression of the zonal flow), it would be a valuable reference for a thorough understanding of the ExB staircase which contains rich local and non-local physics.

I recommend a minor revision by discussing the following issues before being accepted by Phys. Plasmas.

1. What would happen to the ExB staircase when the collisionality is included in the local simulation? It is well-known that residual zonal flow in a tokamak damps by collisional friction between trapped and passing ions [Hinton-Rosenbluth, PPCF '99], and it significantly affects the turbulence-zonal flow dynamics [Lin et al., PRL '99], [Diamond et al., PPCF '05] especially in the absence of nonlinear zonal flow damping [Kim-Diamond, PoP '02].

    > Not goal of ths paper to explain the generation of zonal flows
    > Brief communication has to be short so it was not the focus
    > Arne Weikl 2017 -> Heat flux threshold gets shifted under collision 
      [ ] Include statement of no collision in the "Setup" paragraph

2. What would happen to the ExB staircase when the mean (profile-driven) radial electric field is included in the local simulation? There have been analytic studies of the effect of the mean ExB flow on the zonal flows [Kim-Diamond, PRL '03], [Kagan-Catto, PoP '09], [Hsu-Diamond, PoP '15], but a discussion from the point of view of the ExB staircase would be helpful for readers, as the distinction between the mean and the zonal ExB flows becomes elusive in the ExB staircase problem.

    > From the paper it is not clear what will happen and we do also not know

3. As a related issue, I request a discussion on the zonal modes (structures, i.e., zonal density, etc) in the local simulation study of the ExB staircase, which could be interpreted as a quasi-periodic profile corrugation. The profile corrugation is thought to be an essential symptom of the presence of the ExB staircase [Hahm-Diamond, JKPS '18] with an argument based on the radial force balance, but a discussion based on a point of view of the zonal mode generation would be interesting.

    > No data to following issue was also not the goal of this paper

Referee #2:

This short paper addresses one of the fundamental questions in magnetized plasma turbulence and self-organization. The corrugations (so-called 'staircase' pattern) observed in several simulations pose a challenge of basic understanding. While the paradigm of turbulence suppression by ExB zonal flows has been established since a long time, the appearance of zonal structures, and in particular what determine their radial size, is still unclear.

The main conclusion of the paper is that the radial size of these structures (mesoscale) are, at least in part, due to 'local' physics - which of course does not mean that nonlocal effects are not important.

My main recommendation is that this paper should be accepted, but a few points should be addressed and a few more simulations are required to really make the case more convincing.

1. The zonal structures, once formed, completely quench turbulence in all cases shown here. The end state is thus a pure zonal flow, which persist indefinitely as they should in the collisionless limit. However, the question remains whether the end state depends on the initial condition. In this work, nothing is said about the latter. A proper statistical study would imply an ensemble of simulations with different initial conditions (meaning same equilibrium but different initial perturbations) to address this issue. The statistically meaningful prediction is that of the ensemble average of these runs. This referee strongly advises the authors to run such a set of ensemble runs for at least two most representative cases. This would really make the case much stronger.

    > [ ] Run simulations with different initial conditions (cosine5, noise, default = cosine2) (Florian Rath 2022 as inspiration)

2. The periodicity of the flux-tube implies that the structures tend to have a radial size equal to an integer divisor of the box size. The cases shown all take integer multiples of the initial radial box size. What if non-integer multiples of the radial box size (e.g. 1.5 x , 2.5 x, ...) are considered?

    > [ ] Run simulation with requested box sizes 1.5x1.5, 2.5x2.5
    > Selected box sizes are the mst convenient ones 
    > Not every box is allowed because of the border conditions

3. It would be helpful to plot the radial profiles of R/L_T. Are these profiles peaked at radial positions of zero ExB shearing rate? In the mind of the authors, is the corrugation of temperature gradient the result of what is essentially a turbulence-ExB zonal flows interaction, or is it an essential part of the establishment of the radial structures?

    > [ ] Additional diagnostics (xs_kyzero_dens, xs_kyzero_ene_perp, xs_kyzero_par) for simulation of point 1. of referee #2
    > [ ] Maybe include one sentence in paper
    > Turbulent transport defines R/L_T. When shearing rate is crossing zero a much smaller gradient is needed

4. It looks a bit strange to give a range of pattern size with 4 digits (abstract and throughout the paper).

    > [ ] Remove the last two digits from range