### 2023-05-21 bachelorthesis/01-dedication.tex
#39: Correction by mayself, #37: Whitspace between formula and sign

\thispagestyle{empty}

\begin{center}
    \vspace*{3cm}
    \begin{tikzpicture}
        \clip(0,0)circle[x radius=3cm, y radius=4cm];
        \node{\includegraphics[width=6cm]{Dedication/blacky1_black_white.jpg}};
    \end{tikzpicture}
    \vspace{2cm}\\
    (\textborn) 29.05.2005 - (\textdied) 12.05.2023\\
    \bigskip
    This thesis is dedicated to my cat \textbf{Blacky}\\ which get nearly 19 years old and followed me\\ since I was 5 years through my life.\\You were always a big part of my life and\\ 
    \bigskip
    I will miss you.\\
    \bigskip
    Manuel Lippert\\
    \today
\end{center}