### 2023-05-21 bachelorthesis/04-declaration.tex

\chapter*{Declaration}
\label{chap:declaration}

The author declares that the material presented in this thesis is his own work, unless explicitly stated otherwise. This thesis is based on
\bigskip
\begin{itemize}
    \item [] \bibentry{Lippert2023}
\end{itemize}
\bigskip
and is a further iteration of this publication (brief communication). It provides additional plots and paragraphs that were not necessary to be included in the final publication. The brief communication can be found in Appendix \ref{append:paper}.