### 2023-06-08 bachelorthesis/04-declaration.tex

\chapter*{Declaration}
\label{chap:declaration}

The author declares that the materials presented in this thesis are his own work, unless explicitly stated otherwise. This thesis is based on
\bigskip
\begin{itemize}
    \item [] \bibentry{Lippert2023}
\end{itemize}
\bigskip
and is a further iteration of this publication (brief communication). It provides additional plots and paragraphs that were included in the publication. The brief communication can be found in Appendix \ref{append:paper}.