### 2023-06-11 bachelorthesis/62-data.tex
#39, #36: New date in thesis, csv and readme

\section{Simulation parameter}
\label{append:data}

\begin{center}
    \captionsetup{type=table}
    \begin{tabular}{l | c | l | l l l l l l l l | l}
            & $\rlt$ & box size  & $\Ns$ & $\Nvpar$ & $\Nmu$ & $d_\mathrm{tim}$ & $k\rho^\mathrm{max}$ & $N_\mathrm{mod}$ & $N_x$ & ikx & finit   \\ \hline 
        1   & 6.0    & 1x1       & 12    & 16       & 6      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        2   & 6.0    & 1x1       & 12    & 32       & 6      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        3   & 6.0    & 1x1       & 12    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        4   & 6.0    & 1x1       & 12    & 64       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        5   & 6.0    & 1x1       & 16    & 16       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        6   & 6.0    & 1x1       & 16    & 32       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        7   & 6.0    & 1x1       & 16    & 48       & 6      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        8   & 6.0    & 1x1       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        9   & 6.0    & 1x1       & 16    & 48       & 9      & 0.025            & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        10  & 6.0    & 1x1       & 16    & 48       & 9      & 0.02             & 0.7                  & 11               & 83    & 5   & cosine2 \\ 
        11  & 6.0    & 1x1       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 43    & 5   & cosine2 \\ 
        12  & 6.0    & 1x1       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 63    & 5   & cosine2 \\ 
        13  & 6.0    & 1x1       & 16    & 64       & 6      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        14  & 6.0    & 1x1       & 16    & 64       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\
        15  & 6.0    & 1.5x1.5   & 16    & 48       & 9      & 0.02             & 1.4                  & 31               & 125   & 5   & cosine2 \\
        16  & 6.0    & 2x1       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 10  & cosine2 \\ 
        17  & 6.0    & 2x2       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        18  & 6.0    & 2.5x2.5   & 16    & 48       & 9      & 0.02             & 1.4                  & 51               & 207   & 5   & cosine2 \\ 
        19  & 6.0    & 3x1       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 15  & cosine2 \\ 
        20  & 6.0    & 3x1.5     & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 10  & cosine2 \\ 
        21  & 6.0    & 3x1.5     & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 10  & cosine2 \\ 
        22  & 6.0    & 3x2.5     & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 6   & cosine2 \\ 
        23  & 6.0    & 3x3       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        24  & 6.0    & 3x3       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine5 \\
        25  & 6.0    & 3x3       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & noise   \\
        26  & 6.0    & 3x5       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 3   & cosine2 \\ 
        27  & 6.0    & 4x1       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 20  & cosine2 \\ 
        28  & 6.2    & 2x2       & 16    & 64       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        29  & 6.2    & 3x3       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        30  & 6.3    & 1x1       & 16    & 64       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
        31  & 6.4    & 3x3       & 16    & 48       & 9      & 0.02             & 1.4                  & 21               & 83    & 5   & cosine2 \\ 
    \end{tabular}
    \captionof{table}{Parameters of simulations performed for this thesis}
    \label{tab:parametersSet}
\end{center}

\newpage

\begin{center}
    \captionsetup{type=table}
    \begin{tabular}{l | c | l | r r l | l l | l }
            & $\rlt$ & box size & time    & timestep    & error\_index & stable   & $\nzf$ & backup \\ \hline 
        1   & 6.0    & 1x1      &  6000.0 & 10000       & ~            & False    & 0      & True   \\ 
        2   & 6.0    & 1x1      & 12000.0 & 20000       & ~            & False    & 0      & True   \\ 
        3   & 6.0    & 1x1      &  6000.0 & 10000       & ~            & False    & 0      & True   \\ 
        4   & 6.0    & 1x1      &  6000.0 & 10000       & ~            & False    & 0      & True   \\ 
        5   & 6.0    & 1x1      & 12000.0 & 20000       & ~            & False    & 0      & True   \\ 
        6   & 6.0    & 1x1      & 12000.0 & 20000       & ~            & False    & 0      & True   \\ 
        7   & 6.0    & 1x1      &  6000.0 & 10000       & ~            & False    & 0      & True   \\ 
        8   & 6.0    & 1x1      &  6000.0 & 10000       & ~            & True     & 1      & True   \\ 
        9   & 6.0    & 1x1      &  7125.0 & 10000       & ~            & True     & 1      & True   \\ 
        10  & 6.0    & 1x1      &  6000.0 & 10000       & ~            & True     & 1      & True   \\ 
        11  & 6.0    & 1x1      &  6000.0 & 10000       & ~            & False    & 0      & True   \\ 
        12  & 6.0    & 1x1      &  6000.0 & 10000       & ~            & False    & 0      & True   \\ 
        13  & 6.0    & 1x1      & 12000.0 & 20000       & ~            & True     & 1      & True   \\ 
        14  & 6.0    & 1x1      & 12000.0 & 20000       & ~            & True     & 1      & True   \\
        15  & 6.0    & 1.5x1.5  &  6000.0 & 10000       & ~            & True     & 2      & True   \\
        16  & 6.0    & 2x1      & 18000.0 & 30000       & ~            & True     & 2      & True   \\ 
        17  & 6.0    & 2x2      & 18000.0 & 30000       & ~            & True     & 2      & True   \\ 
        18  & 6.0    & 2.5x2.5  &  3402.1 &  5672       & ~            & True     & 3      & True   \\
        19  & 6.0    & 3x1      & 48000.0 & 80000       & ~            & True     & 3      & True   \\ 
        20  & 6.0    & 3x1.5    & 16840.3 & 28068       & ~            & True     & 4      & True   \\ 
        21  & 6.0    & 3x1.5    & 14170.0 & 23618       & 23618        & True     & 3      & True   \\ 
        22  & 6.0    & 3x2.5    &  6000.0 & 10000       & ~            & True     & 3, 4   & True   \\ 
        23  & 6.0    & 3x3      &  7847.0 & 13083       & ~            & True     & 4      & True   \\ 
        24  & 6.0    & 3x3      &  3454.8 &  5758       & 1840-1845    & True     & 4      & True   \\ 
        25  & 6.0    & 3x3      &  5808.0 &  9680       & 1840-1842    & True     & 3      & True   \\ 
        26  & 6.0    & 3x5      &  4769.0 &  7958       & ~            & True     & 4      & True   \\ 
        27  & 6.0    & 4x1      & 30000.0 & 50000       & 47084        & True     & 4      & True   \\ 
        28  & 6.2    & 2x2      &  6000.0 & 10000       & ~            & True     & 2      & True   \\ 
        29  & 6.2    & 3x3      & 14682.7 & 24475       & 14473-16132  & True     & 3      & True   \\ 
        30  & 6.3    & 1x1      &  6000.0 & 10000       & ~            & False    & 0      & True   \\ 
        31  & 6.4    & 3x3      & 12000.0 & 20000       & ~            & False    & 0      & True   \\ 
    \end{tabular}
    \captionof{table}{Time steps with index which should be exclude in data analysis. Additional the stable column which indicates if turbulence subdued in simulation with the corresponding zonal flow mode number $\nzf$ (0 stands for turbulence not stable) and if the data is saved on the NAS of TPV (backup column)}
    \label{tab:parametersTime}
\end{center}

\newpage

\begin{center}
    \captionsetup{type=table}
    \begin{tabular}{l | l }
            & path                                                              \\ \hline
        1   & ./data/S6\_rlt6.0/boxsize1x1/Ns12/Nvpar16/Nmu6                    \\ 
        2   & ./data/S6\_rlt6.0/boxsize1x1/Ns12/Nvpar32/Nmu6                    \\
        3   & ./data/S6\_rlt6.0/boxsize1x1/Ns12/Nvpar48/Nmu9                    \\ 
        4   & ./data/S6\_rlt6.0/boxsize1x1/Ns12/Nvpar64/Nmu9                    \\ 
        5   & ./data/S6\_rlt6.0/boxsize1x1/Ns16/Nvpar16/Nmu9                    \\ 
        6   & ./data/S6\_rlt6.0/boxsize1x1/Ns16/Nvpar32/Nmu9                    \\ 
        7   & ./data/S6\_rlt6.0/boxsize1x1/Ns16/Nvpar48/Nmu6                    \\ 
        8   & ./data/S6\_rlt6.0/boxsize1x1/Ns16/Nvpar48/Nmu9                    \\ 
        9   & ./data/S6\_rlt6.0/boxsize1x1/Ns16/Nvpar48/Nmu9/dtim0.025          \\ 
        10  & ./data/S6\_rlt6.0/boxsize1x1/Ns16/Nvpar48/Nmu9/krhomax0.70/Nmod11 \\ 
        11  & ./data/S6\_rlt6.0/boxsize1x1/Ns16/Nvpar48/Nmu9/Nx43               \\ 
        12  & ./data/S6\_rlt6.0/boxsize1x1/Ns16/Nvpar48/Nmu9/Nx63               \\ 
        13  & ./data/S6\_rlt6.0/boxsize1x1/Ns16/Nvpar64/Nmu6                    \\ 
        14  & ./data/S6\_rlt6.0/boxsize1x1/Ns16/Nvpar64/Nmu9                    \\
        15  & ./data/S6\_rlt6.0/boxsize1.5x1.5/Ns16/Nvpar48/Nmu9                \\
        16  & ./data/S6\_rlt6.0/boxsize2x1/Ns16/Nvpar48/Nmu9                    \\ 
        17  & ./data/S6\_rlt6.0/boxsize2x2/Ns16/Nvpar48/Nmu9                    \\ 
        18  & ./data/S6\_rlt6.0/boxsize2.5x2.5/Ns16/Nvpar48/Nmu9                \\
        19  & ./data/S6\_rlt6.0/boxsize3x1/Ns16/Nvpar48/Nmu9                    \\ 
        20  & ./data/S6\_rlt6.0/boxsize3x1.5/Ns16/Nvpar48/Nmu9                  \\ 
        21  & ./data/S6\_rlt6.0/boxsize3x1.5/Ns16/Nvpar48/Nmu9/Broken           \\ 
        22  & ./data/S6\_rlt6.0/boxsize3x2.5/Ns16/Nvpar48/Nmu9                  \\ 
        23  & ./data/S6\_rlt6.0/boxsize3x3/Ns16/Nvpar48/Nmu9                    \\ 
        24  & ./data/S6\_rlt6.0/boxsize3x3/Ns16/Nvpar48/Nmu9/cosine5            \\
        25  & ./data/S6\_rlt6.0/boxsize3x3/Ns16/Nvpar48/Nmu9/noise              \\
        26  & ./data/S6\_rlt6.0/boxsize3x5/Ns16/Nvpar48/Nmu9                    \\ 
        27  & ./data/S6\_rlt6.0/boxsize4x1/Ns16/Nvpar48/Nmu9                    \\ 
        28  & ./data/S6\_rlt6.2/boxsize2x2/Ns16/Nvpar64/Nmu9                    \\ 
        29  & ./data/S6\_rlt6.2/boxsize3x3/Ns16/Nvpar48/Nmu9                    \\ 
        30  & ./data/S6\_rlt6.3/boxsize1x1/Ns16/Nvpar64/Nmu9                    \\ 
        31  & ./data/S6\_rlt6.4/boxsize3x3/Ns16/Nvpar48/Nmu9                    \\ 
    \end{tabular}
    \captionof{table}{Data location for each simulation}
    \label{tab:parametersPath}
\end{center}