### 2023-06-16 bachelorthesis/02-acknowledgement.tex
#39: Correction Florian Rath

\chapter*{Acknowledgement}
\label{chap:acknowledgement}

First, I would like to extend my greatest gratitude to my supervisor Prof. Arthur Peeters and Dr. Florian Rath. I would like to thank both of you for your guidance and support during the submission process of the brief communication and the helping hands during the writing process of this thesis and development of the restart script. \bigskip

I would like to thank Bernhard Winkler and Markus Hilt for the great technical support regarding my questions to the \texttt{btrzx1}-cluster which helped me develop the restart script as it is now. \bigskip

Also, I would like to thank my best friend, Paul Schwanitz, for his support in many situations and the good discussions in- and outside of academia. In addition I would like to thank Anna-Maria Pleyer, Dominik Müller and Stefan Barthelmann for the great time of sharing the office together. \bigskip

I thank Anna-Maria Pleyer and my sister Cornelia Lippert for proofreading this thesis and brief communication, which helps me getting better at English spelling. \bigskip

Outside of academia, I would like to extend my gratitude to my parents, my brother, his wife and daughter, my sister and my girlfriend for their encouragement, endurance and support while writing this thesis. 
% I am aware that sometimes I am a workaholic who needs to be reminded about the life outside of my academic work.