### 2023-06-20 pictures/Theory/ExB-Drift.tex

% \usetikzlibrary{mindmap,backgrounds}
% \usetikzlibrary{decorations.pathmorphing}
% \usetikzlibrary{decorations.markings}
% \usetikzlibrary{arrows.meta,bending}

\begin{tikzpicture}
    % Electron
    \draw[thick,decoration={segment length=2mm, amplitude=0.3cm, coil},decorate,<-,blue] (3,0) -- (0,0);
    \draw [fill=blue, blue] (0,0.2) circle (0.2) node[white, font=\boldmath]{$-$};
    % Ion
    \draw[thick,decoration={coil, aspect = -0.5, amplitude = -0.9cm, segment length=6mm},decorate,<-,red] (3,-2) -- (0,-2);
    \draw [fill=red, red] (0,-2.2) circle (0.2) node[white, font=\boldmath]{$+$};
    % Electric field
    \draw[thick,Notablue, ->] (4,-2) -- (4,0) node[above right, Notablue]{$\vect{E}$};
    % Drift velocity
    \draw[thick,Notaorange, ->] (0.5,-0.7) -- (2.5,-0.7) node[right, Notaorange]{$\vect{v}_{E}$};
    % Magnetic field
    \draw [thick, Notagreen] (4,-2.5) circle (0.2);
    \filldraw [Notagreen] (4,-2.5) circle (1.5pt);
    \draw (4.3,-2.2) node[Notagreen]{$\vect{B}$};
\end{tikzpicture}