### 2023-06-20 pictures/Theory/Inhomogenous-Magnetic-Field.tex

% \usetikzlibrary{mindmap,backgrounds}
% \usetikzlibrary{decorations.pathmorphing}
% \usetikzlibrary{decorations.markings}
% \usetikzlibrary{arrows.meta,bending}

{
\def\ang{36}
\def\Rx{0.7}
\def\Ry{1.14}
\def\h{0.5}
\def\H{3}
\def\W{4}
\def\NB{2}

\begin{tikzpicture}

	\coordinate (O) at (0,0);
	\coordinate (N) at (0,0.24*\H);
	\coordinate (M) at (0,0.45*\H);
	\coordinate (B) at (\ang:\H);

	% Magnetic momentum (1/2)
	\draw [thick, Notared] (0,\Ry) arc (90:270:{\Rx} and {\Ry});
	% Magnetic field
	\foreach \i [evaluate={\y=(0.34*\H)*\i^2/\NB; \out=1*\i^2; \in=180+10*\i^2; \f=0.80-0.10*\i;}] in {1,...,\NB}{
	  \draw [thick, Notagreen, <-] (-0.95*\H, 0.7*\y/\H) to[out= \out,in= \in,looseness=0.8] (0.25*\W, \y);
	  \draw [thick, Notagreen, <-] (-0.95*\H,-0.7*\y/\H) to[out=-\out,in=-\in,looseness=0.8] (0.25*\W,-\y);
	}
	\node [Notagreen] at (0.26*\W,0.52*\H) {$\vect{B}$};
	% Force
	\draw [thick, Notagrey, ->] ( 95:{\Rx} and {\Ry}) --++ (-25:0.8*\Ry) node[right=0, Notagrey] {$\vect{F}_{\parallel,\nabla_{\!\parallel} B}$};
	\draw [thick, Notagrey, ->] (-95:{\Rx} and {\Ry}) --++ ( 25:0.8*\Ry) node[right=1, Notagrey] {$\vect{F}_{\parallel,\nabla_{\!\parallel} B}$};
	% Magnetic momentum (2/2)
	\draw [thick, Notared] (0,\Ry) arc (90:-90:{\Rx} and {\Ry});
	\draw [thick, Notared, ->] (0,0) --++ (-0.3*\W,0) node[left, Notared] {$\vect{\mu}$};
	% Particle
	\filldraw [Notablue] (-95:{\Rx} and {\Ry}) circle (2pt);
	\filldraw [Notablue] ( 95:{\Rx} and {\Ry}) circle (2pt);
	\draw (-3,1) node[above right, Notablue]{Particle};

\end{tikzpicture}}