### 2023-06-20 pictures/Theory/Lorentz-Force.tex

% \usetikzlibrary{mindmap,backgrounds}
% \usetikzlibrary{decorations.pathmorphing}
% \usetikzlibrary{decorations.markings}
% \usetikzlibrary{arrows.meta,bending}

\begin{tikzpicture}
    % Circle
    \draw (0,0) circle (1.5);
    \filldraw (0,0) circle (1.5pt);
    % Force
    \draw[thick, Notagrey, <-] (90:0.5) node[right, Notagrey]{$\vect{F}_{\perp, B}$} -- ++ (90:1.0) node[coordinate] (x) {};
    % Larmor radius
    \draw[thick, black, ->] (0,0) -- ++(0:1.5);
    \draw (0.6,-0.4) node {$\rho_\mathrm{L} = \frac{mv_{\perp}}{|q|B}$};
    % Velocity
    \draw[thick,Notaorange, ->] (x) -- ++(180:1cm) node[above, Notaorange]{$\vect{v}$};
    % Cyclontron freqency
    \draw[thick, ->, Notared] (200:1.1*1.5) arc(200:240:1.1*1.5);
    \draw (240:1.4*1.5) node[Notared] {$\omega_\mathrm{c} = \frac{|q|B}{m}$};
    % Particle
    \filldraw [Notablue] (x) circle (2pt);
    \draw (x) node[above right, Notablue]{Particle};
    % Magnetic field
    \draw [thick, Notagreen] (1.5,-1.5) circle (0.2);
    \filldraw [Notagreen] (1.5,-1.5) circle (1.5pt);
    \draw (1.8,-1.2) node[Notagreen]{$\vect{B}$};
\end{tikzpicture}