### 2023-06-20 bachelorthesis/33-cluster.tex

\section{btrzx1 Cluster}
\label{sec:cluster}

The simulation itself will be performed on the \texttt{btrzx1}-cluster of the University Bayreuth. This cluster has a wall time of 24 hours, an total amount of 345 compute nodes with
two AMD Epyc Processors (2nd generation) with 16 cores each and 128 GB of main memory each node and as resource manager \texttt{Slurm}. \cite{btrzx1}

The simulations are performed on the \texttt{/scratch} directory because the \texttt{/home} folder has a disk space limit of \SI{80}{\giga\byte} which is not enough to perform long runs of GKW.