### 2023-06-20 bachelorthesis/50-conclusion.tex

% 5. Closure

\chapter{Conclusion}
\label{chap:close}

\thispagestyle{empty}
\newpage

In this thesis the minimal resolution for simulations with GKW in the Cyclone Base parameter were observed in which the number of grid points for the parallel velocity $\Nvpar$ could be reduced from 64 to 48, which halfed the time until suppression of turbulence. \bigskip

Additionally, the active development of a restart script in \texttt{python3} led to further convenience during the task of performing simulations on the \texttt{btrzx1} cluster. \bigskip

Through careful tests this bachelor thesis confirms the radial size convergence of the $\exb$ staircase pattern in local gyrokinetic flux tube simulations of ion temperature gradient (ITG)-driven turbulence.
A mesoscale pattern size of $\sim 57 - 76\,\rhoth$ is found to be intrinsic to ITG-driven turbulence for Cyclone Base Case parameters.
This length scale is somewhat larger compared to results from global studies with finite $\rho_\ast$, which report of a few $10\,\rhoth$ \cite{Pradalier2010}, and has to be considered the proper mesoscale in the local limit $\rho_\ast \rightarrow 0$.  
The occurrence of this mesoscale implies that non-locality, in terms of Ref. \citenum{Pradalier2010}, is inherent to ITG-driven turbulence, since avalanches are spatially organized by the $\exb$ staircase pattern \cite{McMillan2009, Pradalier2010, Rath2016, Peeters2016}.