### 2023-06-21 presentation/presentation.tex

% Author         : Manuel Lippert
% Version        : 1.1
% Created on     : 11.06.2023
% Last Edited on : 11.06.2023

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                    !!! Compile with LuaLaTex !!!
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[compress,aspectratio=1610,noflama]{beamer}

%--------------------------------------------------------------------------
% Theme
%--------------------------------------------------------------------------

\usepackage{theme/beamerthemefancy}
\usetheme{fancy}

\usepackage{theme/dtklogos} % must be loaded after theme

%--------------------------------------------------------------------------
% Common packages
%--------------------------------------------------------------------------

\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{multimedia}
\usepackage{media9}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{soul}
%\usepackage[numbers, super]{natbib}
\usepackage{tikz}
\usepackage{multirow}

%--------------------------------------------------------------------------
% Variables
%--------------------------------------------------------------------------

\input{../bachelorthesis/colors.tex}
\input{../bachelorthesis/flowdiagramm.tex}
\input{../bachelorthesis/variables.tex}
\input{../bachelorthesis/tikz.tex}

%--------------------------------------------------------------------------
% Variables
%--------------------------------------------------------------------------

\newcommand{\backgroundpicture}{Comparison/Boxsize/S6_rlt6.0_boxsize1-2-3-4x1_Ns16_Nvpar48_Nmu9_wexb_comparison.pdf}
\newcommand{\backgroundxshift}{0.27\paperwidth}
\newcommand{\backgroundwidth}{0.7\paperwidth}

\newcommand{\logopicture}{Uni_Logo_white_black}

%--------------------------------------------------------------------------
% Functions, Graphics, Tikz, Underlines
%--------------------------------------------------------------------------

\graphicspath{{../pictures/}}

\newcommand{\vect}[1]{\boldsymbol{\mathbf{#1}}}

\setul{0.3ex}{0.2ex}

\makeatletter 
\let\UL\ul 
\renewcommand\ul{
	\let\set@color\beamerorig@set@color 
	\let\reset@color\beamerorig@reset@color 
	\UL} 
\makeatother

%--------------------------------------------------------------------------
% Document
%--------------------------------------------------------------------------

\title{Size Convergence of the E$\times$B \\ Staircase Pattern in Flux Tube \\ Simulations of Ion t
Temperature \\ Gradient-Driven Turbulence}
\subtitle{~}
\author{Manuel Lippert}
\institute{Theoretical Physics V}

\begin{document}

	\maketitle

	%\begin{frame}{Gliederung}
	%	\tableofcontents[hideallsubsections]
	%\end{frame}

	\section*{Motivation}
	\begin{frame}
		\frametitle{Motivation}

		\begin{center}
			\begin{itemize}
				\item <2-> Ion temperature gradient-driven turbulence close to marginal stability exhibits Zonal flow pattern formation on mesoscales 
				\item <3-> Patterns are called $\exb$ staircase structures
			\end{itemize}
		\end{center}
		\bigskip
		\onslide<4-> Observed pattern formation in
		\begin{center}
			\begin{itemize}
				\item <5-> local gradient-driven flux-tube simulations, including collisions and background $\exb$ shear,
				\item <6-> local flux-driven realizations, including mean electric field shear,
				\item <7-> global gradient-driven and global flux-driven studies.
			\end{itemize}
			\bigskip
			\onslide<8-> \textbf{Does the basic pattern size always converges to the box size, or is there a typical mesoscale size inherent to staircase structures also in a local flux-tube description?}
		\end{center}
	\end{frame}

	\section*{Theory}

	\begin{frame}
		\frametitle{Charged Particle Motion}

		\begin{center}
			\begin{tabular}{>{\onslide<2->}c<{\onslide}@{\hspace{0.5cm}} >{\onslide<3->}c<{\onslide}@{\hspace{0.5cm}} >{\onslide<4->}c<{\onslide}}
				\textbf{Lorentz force} & 
				\textbf{Electric force} & 
				\textbf{Inhomogeneous magnetic field} \\
				$F_{\perp, B} = |q| v_\perp B$ & 
				$F_{\parallel,E} = qE_\parallel$ & 
				$F_{\parallel,\nabla_{\!\parallel} B} = - \underbrace{\frac{mv^2_{\perp}}{2B}}_{\mu} \nabla_{\!\parallel} B$ \\
				\input{../pictures/Theory/Lorentz-Force.tex} & 
				\input{../pictures/Theory/Electric-Force.tex} & 
				\input{../pictures/Theory/Inhomogenous-Magnetic-Field.tex} \\
			\end{tabular}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Drift in the Gyrocenter}
		\begin{center}
			\begin{tabular}{>{\onslide<2->}c<{\onslide} >{\onslide<3->}c<{\onslide} >{\onslide<3->}c<{\onslide}}
				{\boldmath $\exb$} \textbf{Drift} & 
				{\boldmath $\nabla B$} \textbf{Drift} & 
				\textbf{Curvature Drift} \\
				$\vect{v}_{E} = \frac{\vect{E}\times\vect{B}}{B^2}$ & 
				$\vect{v}_{\nabla B} = \frac{m v^2_{\perp}}{2 q}\frac{\vect{B}\times \nabla B}{B^3}$ & 
				$\vect{v}_{\kappa} = \frac{m v^2_{\parallel}}{qB^2}\vect{B}\times \underbrace{\vect{\kappa}}_{\frac{\nabla B}{B}} = \frac{m v^2_{\parallel}}{q} \frac{\vect{B}\times \nabla B}{B^3}$\\[1cm]
				\input{../pictures/Theory/ExB-Drift.tex} & 
				\multicolumn{2}{>{\onslide<3->}c<{\onslide}}{\input{../pictures/Theory/GradB-Drift.tex}} \\
			\end{tabular}
		\end{center}

	\end{frame}
	\begin{frame}
		\frametitle{Magnetic Confinement in Tokamak}
		\begin{center}
			\begin{tabular}{>{\onslide<2->}c<{\onslide} >{\onslide<2->}c<{\onslide}}
				\includegraphics[width=0.6\paperwidth]{Theory/Tokamak-Torus.pdf} & $\beta = \frac{nT}{\mu_0 B^2/2}$
			\end{tabular}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Ion Temperature Gradient (ITG)-driven Instability}
		\begin{center}
			\onslide<2->\input{../pictures/Theory/ITG-Turbulence_Presentation.tex}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Zonal Flows}
		\begin{center}
			\onslide<2->\includegraphics[width = 0.8\paperwidth]{Theory/Zonal-Flow-Generation.pdf}
			%\begin{itemize}
			%	\item <3-> Zonal flows are linked to the $\exb$ flows
			%	\item <4-> Zonal flow is tangential to the flux-surfaces and does not contribute to the turbulent transport across the flux-surfaces
			%\end{itemize}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Shearing Rate}

		\onslide<2-> Pattern formation occurs in the stabilized tokamak plasma the so-called $\exb$ staircase structure. \\\bigskip

		Properties:
		\begin{enumerate}
			\item[(1)] <3-> They occur on a \textit{radial mesoscale} of order of $\rho_\mathrm{L} < 10\,\rhoth < R$
			\item[(2)] <4-> The structure are \textit{quasi stationary} in space and vary on time scales much longer than the typical turbulence time scales.
			\item[(3)] <5-> They have a \textit{typical amplitude} of the order of $10^{-1} \vth/R$.
			\item[(4)] <6-> Turbulent transport occurs in \textit{avalanches} with propagation strongly linked to the local $\exb$ shearing rate.
		\end{enumerate}
		\onslide<7->{
		\begin{center}
			\begin{tabular}{c c}
				$\wexb = \frac{1}{2} \frac{\partial^2 \langle \phi \rangle}{\partial \xcoord^2}$ & $\langle \phi \rangle = \frac{1}{L_\ycoord} \int_0^{L_\ycoord} \mathrm{d}\ycoord ~ \phi(\xcoord,\ycoord,s=0)$ \\
			\end{tabular}
		\end{center}
		}
	\end{frame}

	%\begin{frame}
	%	\frametitle{Vlasov Equation}
	%	\begin{gather*}
	%		\frac{\partial f}{\partial t} + \frac{\partial f}{\partial \vect{x}} \cdot \frac{\mathrm{d}\vect{x}}{\mathrm{d}t} + \frac{\partial f}{\partial \vect{v}} \cdot \frac{\mathrm{d}\vect{v}}{\mathrm{d}t} = 0
	%	\end{gather*}
	%	\begin{gather*}
	%		n = \int \mathrm{d}\vect{v}\,f(\vect{x}, \vect{v}, t) \qquad j = q \int \mathrm{d}\vect{v}\, v f(\vect{x}, \vect{v}, t)
	%	\end{gather*}
	%	\begin{gather*}
	%		\begin{aligned}
	%			\nabla \cdot \vect{B} &= 0 &\qquad \nabla \times \vect{B} &= \mu_0\left( \sum_\mathrm{species} j + \epsilon_0 \frac{\partial \vect{E}}{\partial t} \right) \\
	%			\nabla \cdot \vect{E} &= \frac{1}{\epsilon_0} \sum_\mathrm{species} qn &\qquad \nabla \times \vect{E} &= - \frac{\partial \vect{B}}{\partial t}
	%		\end{aligned}
	%	\end{gather*}
	%\end{frame}

	\begin{frame}
		\frametitle{Gyrokinetics}

		\only<4>{
			\begin{center}
				\includegraphics[width=0.85\paperwidth]{Theory/Gyro-Center-Coordinates.pdf}		
			\end{center}
		}

		\begin{tabular}{l l l}

			\onslide<2->{$\frac{\partial f}{\partial t} + \frac{\partial f}{\partial \vect{x}} \cdot \frac{\mathrm{d}\vect{x}}{\mathrm{d}t} + \frac{\partial f}{\partial \vect{v}} \cdot \frac{\mathrm{d}\vect{v}}{\mathrm{d}t} = 0$} & 
			\onslide<2->{$f(\vect{x}, \vect{v}, t)$} & 
			\onslide<2->{Vlasov Equation} \\[0.7cm]

			\multicolumn{2}{>{\onslide<3->}l<{\onslide}}{$\Downarrow$ \qquad $\frac{\omega}{\omega_\mathrm{c}} \sim \frac{k_\parallel}{k_\perp} \sim \frac{\rhoth}{L_\mathrm{n}} \sim \frac{\delta n}{n_0} \sim \frac{\delta B}{B_0} \sim \frac{v_\mathrm{d}}{\vth} \sim \epsilon_\mathrm{g}$} & 
			\onslide<3->{Gyrokinetic Ordering} \\[0.7cm]

			\onslide<5->{$\frac{\partial f}{\partial t} + \frac{\partial f}{\partial \vect{X}} \cdot \frac{\mathrm{d}\vect{X}}{\mathrm{d}t} + \frac{\partial f}{\partial v_\parallel} \cdot \frac{\mathrm{d}v_\parallel}{\mathrm{d}t} = 0$} & 
			\onslide<5->{$f(\vect{X}, v_{\parallel}, \mu)~;~\frac{\mathrm{d}\mu}{\mathrm{d}t} = 0$} & 
			\onslide<5->{Gyrokinetic Formalism} \\[0.7cm]

			\multicolumn{2}{>{\onslide<6->}l<{\onslide}}{$\Downarrow$ \qquad $f = f_0 + \delta f$} & 
			\onslide<6->{$\delta f$ Approx \& Local Limit} \\[0.7cm]

			\multicolumn{2}{>{\onslide<7->}l<{\onslide}}{$\frac{\partial g}{\partial t} + \vect{v}_\chi \cdot \nabla g + (v_\parallel \vect{b} + \vect{v}_\mathrm{D}) \cdot \nabla(\delta f) - \frac{\mu B}{m}\frac{\vect{B}\cdot\nabla B}{B^2}\frac{\partial (\delta f)}{\partial v_\parallel} = S$} &
			\onslide<7->{Gyrokinetic Equation} \\[0.7cm]
			\multicolumn{3}{>{\onslide<7->}l<{\onslide}}{$g \sim (f_0 + \delta f)\qquad S \sim (f + f_0)\qquad \vect{v}_\mathrm{D} \sim \nabla B \qquad \vect{v}_\chi \sim \exb \qquad f(\xcoord, \ycoord, s, v_\parallel, \mu)$} \\
		\end{tabular}
	\end{frame}

	\section*{Material and Methods}

	\begin{frame}
		\frametitle{Simulation Setup}

		\begin{center}
			\begin{itemize}
				\item <2-> Non-linear flux tube version of Gyrokinetic Workshop (GKW) 
				\item <3-> Circular concentric flux surfaces 
				\item <4-> Standard resolution with 6th order (S6) 
				\item <5-> Cyclone Base Case (CBC) parameters $\rlt = 6.0$ 
				\item <6-> Turbulence level is quantified by the turbulent heat conduction coefficient $\chi$ (normalized by $\rhoth^2 \vth/R$ ($\vth = \sqrt{2 T/m}$)) 
				\item <7-> Quantities $\rhoth$, $R$, $T$, $\vth$ and $m$ are referenced quantities
				\item <8-> Standard box size $(L_\xcoord,~L_\ycoord) = (76.3,~89.8)\,\rhoth$
			\end{itemize}
			\onslide<9->{
			\bigskip
			\begin{tabular}{l | ccccc | ccccc | c | cc}
				& $N_\mathrm{m}$ & $N_x$ & $N_\mathrm{s}$ & $N_{\nu_\parallel}$ & $N_\mu$ & $D$ & $\nu_\mathrm{d}$           & $D_{\nu_\parallel}$ & $D_x$ & $D_y$ & Order & $k_y\rho$ & $k_x\rho$ \\
				\hline
				S6   & 21    & 83    & 16    & 64                  & 9       & 1   & $|\nu_\parallel|$ & 0.2                 & 0.1   & 0.1   & 6     & 1.4       & 2.1       \\
			\end{tabular}}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Diagnostics}

		\begin{center}
			\onslide<2-> $\wexb = \sum_{\kzf} \hatwexb(\kzf,t) \, \exp(\mathrm{i} \kzf \xcoord)$ \\[0.5cm]
			\onslide<3-> $\kzf = 2\pi \nzf/L_\xcoord$ \\[0.5cm]
			\onslide<4-> $\hatwexbamp = 2 |\hatwexb(\kzf,t)|$ \\[0.5cm]
			\begin{itemize}
				\item <5-> Zonal flow mode that dominates the $\exb$ staircase pattern are called \textbf{basic mode} 
				\item <6-> The basic mode exhibits the maximum amplitude in the spectrum $\hatwexbamp$
			\end{itemize}
		\end{center}
	\end{frame}
	\begin{frame}
		\frametitle{Restart Script}
		\pause

		\newcommand{\distance}{2.5cm}

		\begin{center}
		    \begin{tikzpicture}[node distance=1.5cm]

		        \node (init)         [init]                                                 {Start / Initialize};
		        \node (startLoop)    [startstop, below of = init]                           {Loop};
		        \node (checkStatus)  [process, below of=startLoop]                          {Check Status};
		        \node (checkData)    [process, below of=checkStatus, yshift = -1.5cm]       {Check Data};
		        \node (checkPending) [decision, below of=checkStatus, xshift=\distance]     {Pending};
		        \node (checkRunning) [decision, below of=checkStatus, xshift=-\distance]    {Running};

		        \draw [arrow] (init)           --                                           (startLoop);
		        \draw [arrow] (startLoop)      --                                           (checkStatus);
		        \draw [arrow] (checkStatus)    -|                                           (checkRunning);
		        \draw [arrow] (checkRunning)   -- node[anchor=south] {No}                   (checkPending);
		        \draw [arrow] (checkPending)   -- node[anchor=north west] {Yes} + ( 3,0) |- (startLoop);
		        \draw [arrow] (checkRunning)   -- node[anchor=north east] {Yes} + (-3,0) |- (startLoop);
		        \draw [arrow] (checkPending)   |- node[anchor=north west] {No}              (checkData);

		    \end{tikzpicture}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Restart Script}

		\newcommand{\distance}{2.5cm}

		\begin{center}
		    \begin{tikzpicture}[node distance=2cm]

		        \node (checkData)    [process]       										{Check Data};
		        \node (error)        [decision, below of=checkData, xshift=-\distance]      {Errors};
		        \node (checkPoint)   [decision, below of=checkData, xshift=\distance]       {Checkpoint};
		        \node (backup)       [process,  below of=error, xshift=-\distance]          {Backup Data};
		        \node (restore)      [process,  below of=checkPoint, xshift=\distance]      {Restore Data};
		        \node (reset)        [process,  below of=checkData, yshift=-2cm]            {Reset Data};
		        \node (checkStep)    [decision, below of = reset]                           {Enough Timesteps};

		        \draw [arrow] (checkData)      -|                                           (error);
		        \draw [arrow] (error)          -- node[anchor=south] {Yes}                  (checkPoint);
		        \draw [arrow] (error)          -| node[anchor=south east] {No}              (backup);
		        \draw [arrow] (checkPoint)     -| node[anchor=south west] {No}              (restore);
		        \draw [arrow] (checkPoint)     |- node[anchor=north west] {Yes}             (reset);
		        \draw [arrow] (backup)         --                                           (checkStep);
		        \draw [arrow] (reset)          --                                           (checkStep);
		        \draw [arrow] (restore)        --                                           (checkStep);

		    \end{tikzpicture}
		\end{center}

	\end{frame}

	\begin{frame}
		\frametitle{Restart Script}

		\newcommand{\distance}{2.5cm}

		\begin{center}
		    \begin{tikzpicture}[node distance=1.5cm]

				\node (init)         [init]                                                 {Start / Initialize};
		        \node (startLoop)    [startstop, below of = init]                           {Loop};
		        \node (checkStatus)  [process, below of=startLoop]                          {...};
		        \node (checkStep)    [decision, below of = checkStatus]                     {Enough Timesteps};
		        \node (end)          [startstop, below of=checkStep, xshift=4cm]            {End};
		        \node (restart)      [init, below of=checkStep, xshift=-4cm]                {Start/Restart};

		        \draw [arrow] (init)           --                                           (startLoop);
		        \draw [arrow] (startLoop)      --                                           (checkStatus);
		        \draw [arrow] (checkStep)      -| node[anchor=south east] {No}              (restart);
		        \draw [arrow] (checkStep)      -| node[anchor=south west] {Yes}             (end);
		        \draw [arrow] (restart)        -- + (-3,0) |-                               (startLoop);
		        \draw [arrow] (end)            -- + ( 3,0) |-                               (init);

		    \end{tikzpicture}
		\end{center}

	\end{frame}

	\section*{Results}

	\begin{frame}
		\frametitle{Variation of Computational Resolution}
		\pause
		\textbf{Goals:}
		\begin{itemize}
			\item Estimate the minimal resolution without numerical dissipation
			\item Reduce \textbf{calculation time} and \textbf{costs} of the simulation
		\end{itemize}
		\pause
		\textbf{Criteria:}
		\begin{enumerate}
			\item[\textbf{(1)}] Subdued turbulence after \textbf{short} time periods
			\item[\textbf{(2)}] Stability for \textbf{long} time periods 
		\end{enumerate}
		\pause
		\textbf{Verification:}
		\begin{enumerate}
			\item Reduce only one number of grid points and look if criterias \textbf{(1), (2)} are satisfied
			\item Reduce to known the minimum number of grid points to verify result in general.
		\end{enumerate}
	\end{frame}

	\begin{frame}
		\frametitle{Benchmark}
		\begin{center}
			\only<2>{\includegraphics[width = 0.8\paperwidth]{S6_rlt6.0/boxsize1x1/Ns16/Nvpar64/Nmu9/S6_rlt6.0_boxsize1x1_Ns16_Nvpar64_Nmu9_eflux.pdf}}

			\only<3>{\includegraphics[width = 0.8\paperwidth]{Comparison/Gradient-Length/S6_rlt6.0-6.3_boxsize1x1_Ns16_Nvpar64_Nmu9_eflux_comparison.pdf}}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Reduction of Grid Points}

		\begin{center}
			\only<2>{\includegraphics[width = 0.8\paperwidth]{Comparison/Resolution/Nvpar/S6_rlt6.0_boxsize1x1_Ns16_Nvpar16-32-48-64_Nmu9_eflux_comparison.pdf}}

			\only<3>{\includegraphics[width = 0.8\paperwidth]{Comparison/Resolution/Nmu/S6_rlt6.0_boxsize1x1_Ns16_Nvpar64_Nmu6-9_eflux_comparison.pdf}}

			\only<4>{\includegraphics[width = 0.8\paperwidth]{Comparison/Resolution/Nmu/S6_rlt6.0_boxsize1x1_Ns16_Nvpar48_Nmu6-9_eflux_comparison.pdf}}

			\only<5>{\includegraphics[width = 0.8\paperwidth]{Comparison/Resolution/Ns/S6_rlt6.0_boxsize1x1_Ns12-16_Nvpar64_Nmu9_eflux_comparison.pdf}}

			\only<6>{
				\textbf{Final Resolution}\\
				\bigskip
				\begin{tabular}{l | ccccc | ccccc | c | cc}
					& $N_\mathrm{m}$ & $N_x$ & $N_\mathrm{s}$ & $N_{\nu_\parallel}$ & $N_\mu$ & $D$ & $\nu_\mathrm{d}$           & $D_{\nu_\parallel}$ & $D_x$ & $D_y$ & Order & $k_y\rho$ & $k_x\rho$ \\
					\hline
					S6   & 21    & 83    & 16    & 48                  & 9       & 1   & $|\nu_\parallel|$ & 0.2                 & 0.1   & 0.1   & 6     & 1.4       & 2.1       \\
				\end{tabular}
			}
		\end{center}
	\end{frame}

	\begin{frame}
		\frametitle{Size Convergence of $\exb$ Staircase Pattern}
		\begin{center}
			\only<2>{\textbf{(1) Radial} \\[0.5cm] $\NR \times \NB \in [1\times1,~2\times1,~3\times1,~4\times1]$}

			\only<3>{\includegraphics[width = 0.8\paperwidth]{Comparison/Boxsize/S6_rlt6.0_boxsize1-2-3-4x1_Ns16_Nvpar48_Nmu9_comparison_thesis.pdf}}

			\only<4>{\includegraphics[height = 0.5\paperheight]{Comparison/Boxsize/S6_rlt6.0_boxsize1-2-3-4x1_Ns16_Nvpar48_Nmu9_wexb_comparison.pdf}}

			\only<5>{\textbf{(2) Isotropic} \\[0.5cm] $\NR \times \NB \in [1\times1,~1.5\times1.5,~2\times2,~2.5\times2.5,~3\times3]$}

			\only<6>{\includegraphics[width = 0.8\paperwidth]{Comparison/Boxsize/S6_rlt6.0_boxsize1x1-2x2-3x3_Ns16_Nvpar48_Nmu9_comparison_thesis.pdf}}

			\only<7>{\includegraphics[height = 0.5\paperheight]{Comparison/Boxsize/S6_rlt6.0_boxsize1x1-2x2-3x3_Ns16_Nvpar48_Nmu9_wexb_comparison.pdf}}

			\only<8>{\includegraphics[height = 0.5\paperheight]{Comparison/Boxsize/S6_rlt6.0_boxsize3x3_Ns16_Nvpar48_Nmu9_cosine2-cosine5-noise_wexb_comparison.pdf}}

			\only<9>{\includegraphics[height = 0.7\paperheight]{S6_rlt6.0/boxsize3x3/Ns16/Nvpar48/Nmu9/noise/S6_rlt6.0_boxsize3x3_Ns16_Nvpar48_Nmu9_noise_wexb_dens_ene_selection.pdf}}

			\only<10>{\textbf{(3) Binormal} \\[0.5cm] $\NR \times \NB \in [3\times1.5,~3\times2.5,~3\times3,~3\times5]$}

			\only<11>{\includegraphics[width = 0.8\paperwidth]{Comparison/Boxsize/S6_rlt6.0_boxsize3x1-1.5-2.5-3-5_Ns16_Nvpar48_Nmu9_comparison_thesis.pdf}}

			\only<12>{\includegraphics[height = 0.5\paperheight]{Comparison/Boxsize/S6_rlt6.0_boxsize3x1-1.5-2.5-3-5_Ns16_Nvpar48_Nmu9_wexb_comparison.pdf}}
		\end{center}

		\only<13>{
			\begin{center}
				$\Rightarrow$ Mesoscale pattern size of: \\[0.3cm]
				$\boxed{\sim 57 - 76\,\rhoth}$ \bigskip
				\begin{itemize}
					\item Non-locality is inherent to ITG-driven turbulence
					\item Avalanches are spatially organized by the E × B staircase pattern
				\end{itemize}
			\end{center}
		}
	\end{frame}

	\begin{frame}
		\frametitle{The finite Heat Flux Threshold}
		\begin{center}
			\onslide<2->\includegraphics[width = 0.8\paperwidth]{Comparison/Gradient-Length/S6_rlt6.0-6.2-6.4_boxsize3x3_Ns16_Nvpar48_Nmu9_eflux_comparison.pdf}

			\onslide<3->$\Rightarrow~\boxed{\rlt|_\mathrm{finite} = 6.3 \pm 0.1}$
		\end{center}
	\end{frame}

	\section*{Conclusion}

	\begin{frame}
		\frametitle{Conclusion}

		\begin{minipage}{0.4\paperwidth}\raggedleft
			\begin{itemize}
				\item <2-> Parallel velocity $\Nvpar$ could be reduced from 64 to 48, which halfed the time until suppression of turbulence
				\item <3-> Restart Script with \texttt{python} led to further convenience during the task of performing simulations
				\item <4-> Mesoscale pattern size of $\sim 57 - 76\,\rhoth$ is found to be intrinsic to ITG-driven turbulence for Cyclone Base Case parameters
				\item <5-> Finite heat flux threshold is located at $\rlt|_\mathrm{finite} = 6.3 \pm 0.1$
			\end{itemize}
		\end{minipage}
		\begin{minipage}{0.4\paperwidth}
			\onslide<4-> \includegraphics[width = 0.5\paperwidth]{Comparison/Boxsize/S6_rlt6.0_boxsize1-2-3-4x1-1.5-2-2.5-3-5_Ns16_Nvpar48_Nmu9_wexb_comparison_thesis.pdf}
		\end{minipage}
	\end{frame}

\end{document}