### 2023-06-23 pictures/Theory/Lorentz-Force.tex
#37, #39: Redo circular motion picture and some grammar

% \usetikzlibrary{mindmap,backgrounds}
% \usetikzlibrary{decorations.pathmorphing}
% \usetikzlibrary{decorations.markings}
% \usetikzlibrary{arrows.meta,bending}

\begin{tikzpicture}
    % Circle
    \draw[lightgray] (0,0) circle (1.5);
    \filldraw (0,0) circle (1.5pt);
    % Force
    \draw[thick, Notagrey, <-] (90:0.5) node[right, Notagrey]{$\vect{F}_{\perp, B}$} -- ++ (90:1.0) node[coordinate] (x) {};
    % Larmor radius
    \draw[thick, black, ->] (0,0) -- ++(0:1.5);
    \draw (0.6,-0.4) node {$\rho_\mathrm{L} = \frac{mv_{\perp}}{|q|B}$};
    % Velocity
    \draw[thick,Notaorange, ->] (x) -- ++(180:1cm) node[left, Notaorange]{$\vect{v}^-$};
    \draw[thick,Notaorange, ->] (0,-1.5) -- ++(180:1cm) node[left, Notaorange]{$\vect{v}^+$};
    % Cyclontron freqency
    \draw[thick, ->, blue] (90:1.0*1.5) arc(90:150:1.0*1.5);
    \draw[thick, <-, red] (210:1.0*1.5) arc(210:270:1.0*1.5);
    \draw (50:1.5*1.5) node[black] {$\omega_\mathrm{c} = \frac{|q|B}{m}$};
    % Charges
    \draw [fill=blue, blue] (x) circle (0.2) node[white, font=\boldmath]{$-$};
    \draw [fill=red, red] (0,-1.5) circle (0.2) node[white, font=\boldmath]{$+$};
    % Magnetic field
    \draw [thick, Notagreen] (1.5,-1.5) circle (0.2);
    \filldraw [Notagreen] (1.5,-1.5) circle (1.5pt);
    \draw (1.8,-1.2) node[Notagreen]{$\vect{B}$};
\end{tikzpicture}